---
title: "Collecting primary data"
output: 
  workflowr::wflow_html:
    code_folding: hide 
editor_options:
  chunk_output_type: console
  code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=F}
library(tidyverse)
library(readxl)
library(lme4)
library(car)
library(lmerTest)
library(emmeans)
library(survival)
library(coxme)
library(kableExtra)
library(compute.es)

# Function to parse results of compute.es::mes(), fes(), etc.
parse_es <- function(x, study, trait, generations, organism, interpretation){
  x %>% 
    as.list() %>% enframe() %>% unnest(value) %>% 
    filter(name %in% c("d", "var.d", "l.d", "u.d")) %>% 
    mutate(type = c("Estimate", "Var", "Lower_95_CI", "Upper_95_CI")) %>% 
    select(-name) %>% 
    pivot_wider(names_from = type, values_from = value) %>% 
    mutate(Study = study, Trait = trait, Generations = generations, 
           Organism = organism, Interpretation = interpretation, 
           Effect_size_source = "Calculated from sample means (pseudoreplicated)",
           .before = 1)
}

# Function to calculate Cohen's d effect size from a 2x2 contingency table
# It first calculates the log odds ratio, then converts this to Cohen's d via logOR * sqrt(3) / pi
# SE of the LOR is calculated using Woolf's method
get_effect_size_2x2_matrix <- function(x){
  
  # calculate log odds ratio as log(a*d / b*c)
  logOR <- as.numeric(log((x[1,1] * x[2,2]) / 
                            (x[2,1] * x[1,2])))
  
  # Standard error of the log OR (Woolf method)
  SE_logOR <- 1/x[1,1] + 1/x[1,2] + 1/x[2,1] + 1/x[2,2]
  
  # 95% CIs of the log OR
  CI_logOR <- logOR + c(-1, 1) * 1.96 *  SE_logOR
  
  # Convert LOR to d:
  d <- logOR * sqrt(3) / pi
  SE_d <- SE_logOR * sqrt(3) / pi
  CI_d <- CI_logOR * sqrt(3) / pi
  
  tibble(Estimate = d,
         Var = SE_d ^ 2,
         Lower_95_CI = CI_d[1],
         Upper_95_CI = CI_d[2])
}
```

# Overview

## Cohen's $d$

This document describes how I estimated Cohen's $d$ for each primary study in the meta-analysis. Cohen's $d$ is a standardised measure of the effect size, defined as the difference between two means, expressed in standard deviations of the variable of interest. Conventionally, Cohen's $d > |0.5|$ is considered a 'large' effect. 

## Meaning of effect size sign

Crucially, effect sizes have a sign (positive or negative, depending on whether mean 1 or mean 2 is the larger) as well as a magnitude. Unlike the magnitude, the sign is determined solely by the researcher: it depends on which of the two means is designated as the reference for comparison.

To ensure the effect size sign is meaningful and comparable across studies, __I defined a positive effect size to indicate that the two means differ in the direction that one would predict if intralocus sexual conflict is present__, and a negative effect to mean that the means differ in a direction opposite to this prediction. 

For example, for a study like Prasad et al. (2007) that removed selection on females and then measured mean male and female fitness relative to a control, we would predict male-evolved populations to have _higher_ male fitness and _lower_ female fitness if intralocus sexual conflict is present. So, a positive effect size should indicate higher male fitness but lower female fitness in the experimental group relative to the control. This is potentially confusing, but it ensures that finding the mean effect size via meta-analysis is useful: a positive mean effect size would mean that experimental evolution studies tend to produce the results predicted if IaSC is common.

To clarify the meaning of each effect size, I provide a column called 'interpretation' in the final table of effect sizes, and document the more confusing/ambiguous decisions in this document below.

## How Cohen's $d$ was calculated

### Preferred method: from the raw data

If the primary study's raw data were available (e.g. on Dryad or Figshare, or as a table in the original paper), I calculated the effect size from an appropriate statistical model that accounted for the structure of the experimental design (e.g. the use of replicate experimental evolution lines), block effects, etc. (following the model used by the original authors wherever possible). The models chosen are documented in the code below.

For response variables where it was appropriate to use Gaussian linear (mixed) models, I first transformed the response variable to have a mean of zero and variance of 1, such that the difference between treatment means estimates by the statistical model is equal to Cohen's d. 

For generalised (e.g. binomial) models that estimated treatment means as a log odds ratio ($\log(\mathrm{OR})$), or a log hazard ratio (for survival models), I converted this effect size to Cohen's d using the standard approximation:

$\log(\mathrm{OR}) = \frac{\sqrt{3}}{\pi}$

For studies where the original data represent counts of binary events tallied up in a $2\times2$ contingency table (e.g. number of surviving and non-surviving individuals from the treatment and control groups) with elements $a$,$b$,$c$ and $d$, I estimated the log odds ratio as $log(a\times d / b\times c)$ and found the SE of the log odds ratio using Woolf's method, i.e. 

\[SE\!\left(\log(\mathrm{OR})\right)
=\sqrt{\frac{1}{a}+\frac{1}{b}+\frac{1}{c}+\frac{1}{d}}\]

### Fall-back method: from the group means

For many studies, the raw data were not available, but the paper provided numerical or graphical information about 1) the means of each treatment group, 2) either the group standard deviations or the standard errors of the means (which allow calculation of the standard deviation as $SD = SE\times\sqrt{N}$), and 3) the sample size N. From these three quantities, one can estimate Cohen's d (I used the function `mes()` from the `compute.es` package). Where possible I obtained these numbers from tables or the main text, but often the data were only contained in figures (e.g. bar charts showing the mean and standard errors). In the latter case, I measured the means and standard errors using the web application [PlotDigitizer](https://plotdigitizer.com/) - this data extraction step is expected to add inaccuracy, but not to add consistent bias.

Using group means is a less accurate method because it is not possible to account for the structure of the experimental design, incurring pseudoreplication. For example, in studies where the authors had 4 replicate experimental evolution lines for 2 treatments and measured 20 individuals per line, the sample size is not really 80 per treatment, since there are 20 replicate measurements of only 8 lines. This will make the effect size measurements appear more precise than they should be, giving extra weight to those studies in the meta-analysis. 

### Workflow in the code below

For each publication, I either load or type out the raw data, display it (in cases where I extracted the raw data or group means from the primary paper), calculate effect size, and then save it to a small .csv file along with meta-data like the paper of origin, the trait being measured, the number of generations of selection prior to measuring the trait, and a plain English interpretation of the meaning of the effect size (e.g. that the trait is higher/lower in one treatment group compared to another such group).

# Sex-limited evolution experiments

This section covers publications that evolved genomic elements (e.g. small regions or whole chromosomes) mostly/entirely within one sex for multiple generations. In general, I predict that this treatment will increase fitness in that sex, reduce fitness in the other sex, and cause phenotypic evolution towards the typical trait values for the evolving sex (e.g. in _Drosophila_, chromosomes that had undergone male-limited evolution in should encode smaller body size and slower development time). The effect size signs selected reflect these expectations.

## Rice (1992)

Rice, W. R. (1992). Sexually antagonistic genes: experimental evidence. _Science_, 256, 1436-1439.

### Rice's original data

These data are shown in Table 1. Letters R, P, S and O denote the 4 eye colours of males that females might mate to, where O (orange-eyed) represents non-focal males and R/P/S are the focal male (which differ in how many copies of the evolved genomic region is carried). The focal males carry alleles from either the control or the population undergoing female-limited evolution of a small genomic region. The R,P,S, and O columns are percentages, with the last column giving the total number of females tested. 

```{r}
df <- tibble(
  replicate = c(1,2,1,2),
  treatment = c("control", "control", "experimental", "experimental"),
  R = c(42, 36, 40, 17),
  P = c(26, 21, 20, 40),
  S = c(22, 33, 22, 26),
  O = c(10, 9, 17, 17),
  total = c(90, 75, 90, 70)
)

df %>% 
  kable() %>% 
  kable_styling()
```

### Calculating effect size

```{r}
# first convert the 4 percentage columns to counts of females (with rounding)
df$R <- round((df$R/100) * df$total)
df$P <- round((df$P/100) * df$total)
df$S <- round((df$S/100) * df$total)
df$O <- round((df$O/100) * df$total)

# Pool the females over the 3 types representing focal males
focal_male_chosen <- rowSums(df %>% select(R, P, S))

# Make a 2x2 contingency table giving number of times the focal vs non-focal male mated, for control vs experimental
x <- matrix(c(focal_male_chosen[1] + focal_male_chosen[2], 
         focal_male_chosen[3] + focal_male_chosen[4], 
         df$O[1] + df$O[2], 
         df$O[3] + df$O[4]), ncol = 2)

get_effect_size_2x2_matrix(x) %>% 
    mutate(Study = "Rice (1992) Science 256, 1436-1439", 
           Trait = "Male fitness", 
           Generations = "29",
           Organism = "Drosophila melanogaster",
           Interpretation = "A positive effect size means that males carrying a genomic region that underwent female-limited evolution had lower fitness compared to males carrying control genomes.",
           Effect_size_source = "Chi-square test on total counts (pseudoreplicated)",
           .before = 1) %>% 
  write_csv("effect_sizes/rice92.csv")
```


## Rice (1996)

Rice, W. R. (1996). Sexually antagonistic male adaptation triggered by experimental arrest of female evolution. _Nature_, 381, 232–234.

### Male fitness

#### Rice's original data

The means for the male fitness were recorded from Figure 1, using PlotDigitizer. 

```{r}
# These data come from using PlotDigitizer web app on data in the first 2 columns of Figure 2
rice96 <- tibble(generation = c(rep(31,4), rep(32, 4)),
       replicate = rep(c("A", "A", "B", "B"), 2),
       treatment = rep(c("ML", "Control", "ML", "Control"), 2),
       fitness = c(1.312563038, 1.040288341, 1.240486316, 0.959122165,
                   1.233782098, 1.018216389, 1.182082107, 0.982122656),
       sample_size = rep(c(4,8), each=4)) 

rice96 %>% 
  mutate(fitness = round(fitness,2)) %>% 
  rename(`Male fitness` = fitness) %>% 
  kable() %>% kable_styling(full_width = F)
```

#### Calculating effect size

Unfortunately the study provides means but no standard error or standard deviation, which we need to compute effect size variance and confidence limits. 

To get an estimate of effect size, we assume here the standard deviation (SD) is the same as the SD in  similar male fitness assays carried out in Wong & Holman (2023; _Evolution_). That study's male fitness assay methods were quite similar, and focused on North American _D. melanogaster_ as well. We load that study's raw data, and find the average standard deviation in proportion of offspring sired relative to competitor males. We then use simulation to generate 10,000 possible datasets for Rice (1996) by generating random Normally distributed numbers using the means from Rice's Figure 1 and SD for Wong & Holman, and from these random datasets calculate 10,000 estimates of the standardised effect size Cohen's d comparing the two treatments. From these 10,000 estimates we also obtain the effect size variance and 95% CIs. 


```{r}
# Rice's description of the male fitness trait:
# Extra vials were constructed by mixing EA, EB, CA', and CB' males with target females (4 and 8 vials each in generations 31 and 32, respectively) using a procedure matching that of the adapting-male compartment of Fig. 1, except that 75% of the males were replaced by competitor males that were homozygous for the marker pP (previously recombined into the genetic background of the LH source stock). Offspring from competitor males were pink-eyed and could be distinguished from red-eyed males carrying the genomic haplotype of the experimental (EA or EB) or control (CA' or CB') males. The ratio of red-eyed to pink-eyed adult male offspring, after adjusting for Mendelian segregation, measured net relative fitness.

# Find the standard deviation in proportion offspring sired in a different study, Wong & Holman 2023 Evolution, for 810 male fitness assays covering 125 fly genotypes
# Then take the average SD across those 125 genotypes, and assume this is representative of the SD in Rice's study
estimated_SD <- read_csv(
  "input_data/male_fitness_raw_data_wong_and_holman.csv", 
  show_col_types = FALSE) %>% 
  mutate(fitness = `Early male focal` / 
           (`Early male focal` + `Early male rival`)) %>% 
  group_by(Line) %>% 
  summarise(mean_fitness = mean(fitness),
            sd_fitness = sd(fitness)) %>% 
  summarise(x = mean(sd_fitness, na.rm=T)) %>% pull(x)


# make n random normal fitness values given mean and SD
make_data <- function(mean_fitness, n, sd){
  rnorm(n, mean_fitness, sd)
}

bootstrapped_effect_sizes <- lapply(1:1000, function(i){
  
  possible_dataset <- map_df(1:8, ~ tibble(
    generation = rice96$generation[.x],
    replicate = rice96$replicate[.x],
    treatment = rice96$treatment[.x],
    fitness = rnorm(n = rice96$sample_size[.x], 
                    mean = rice96$fitness[.x], 
                    sd = estimated_SD))) %>% 
    mutate(line = paste(treatment, replicate, sep = "_"),
           fitness = as.numeric(scale(fitness))) # scale fitness
  
  # do a linear model (effect size of treatment is equal to Cohen's d)
  as.numeric(summary(lm(fitness ~ treatment + line, data = possible_dataset))$coefficients[2,1:2])
}) %>% 
  do.call("rbind", .) %>% 
  as.data.frame() %>% as_tibble() %>% 
  mutate(V2 = V2 ^ 2) %>% # Square the effect size SE to get the variance
  rename(Estimate = V1, Var = V2)

bootstrapped_effect_sizes %>% 
  summarise(Estimate = mean(Estimate), # bootstrapped Cohen's d effect size
            Var = mean(Var)) %>%       # bootstrapped variance in Cohen's d effect size
  mutate(Lower_95_CI = Estimate - 1.96 * sqrt(Var),
         Upper_95_CI = Estimate + 1.96 * sqrt(Var)) %>% 
    mutate(Study = "Rice (1996) Nature 381, 232–234", 
           Trait = "Male fitness", 
           Generations = "32",
           Organism = "Drosophila melanogaster",
           Interpretation = "The positive effect size means that males carrying a genomic region that underwent male-limited evolution had higher fitness than males carrying control genomes",
           Effect_size_source = "Calculated from means, with simulation used to estimate sampling variance (see archived analysis code)",
           .before = 1)  %>% 
  write_csv("effect_sizes/rice96.csv")
```


## Rice (1998)

Rice, W. R. (1998). Male fitness increases when females are eliminated from the gene pool: Implications for the Y chromosome. _PNAS_, 95, 6217–6221.

### Mating speed  

Data obtained from Figure 1A, using PlotDigitiser. Rice placed males with male-evolved or control genomes in vials with virgin females, and counted the number of mating pairs after 15 minutes. This is a rough measure of how quickly males induce virgin females to mate and presumably correlates with males' ability to find/court females, attractiveness and mating duration. 

#### Rice's original data

```{r}
x <- matrix(c(52, 3, 42, 13), nrow = 2)
colnames(x) <- c("male-limited", "control")
rownames(x) <- c("mated", "unmated")
kable(x) %>% kable_styling(full_width = F)
```

#### Calculating effect size

```{r}
get_effect_size_2x2_matrix(x) %>% 
    mutate(Study = "Rice (1998) PNAS, 95, 6217–6221", 
           Trait = "Male mating speed", 
           Generations = "38",
           Organism = "Drosophila melanogaster",
           Interpretation = "The positive effect size means that males carrying a haplotype that underwent male-limited evolution mated with previously-unmated females faster than males carrying control haplotypes",
           Effect_size_source = "Chi-square test on total counts (pseudoreplicated)",
           .before = 1) %>% 
  write_csv("effect_sizes/rice98_1.csv")
```

### Male attractiveness to mated females

This data is from Figure 1B (early measure), and shows the number of previously mated female flies mating vs not-mating with either EAB or CAB males. More females remating implies that the male type is better able to achieve matings with mated females. Females remated more with EAB males than CAB males.

#### Rice's raw data
```{r}
x <- matrix(c(54, 25, 140-54, 140-25), nrow=2)
colnames(x) <- c("n_remating", "n_not_remating")
rownames(x) <- c("EAB", "CAB")
kable(x) %>% kable_styling(full_width = F)
```

#### Effect size
```{r}
get_effect_size_2x2_matrix(x) %>% 
    mutate(Study = "Rice (1998) PNAS, 95, 6217–6221", 
           Trait = "Male attractiveness", 
           Generations = "38",
           Organism = "Drosophila melanogaster",
           Interpretation = "The positive effect size means that males carrying a haplotype that underwent male-limited evolution attained more matings with previously-mated females compared to males carrying control haplotypes",
           Effect_size_source = "Chi-square test on total counts (pseudoreplicated)",
           .before = 1) %>% 
  write_csv("effect_sizes/rice98_2.csv")
```

### Sperm defence

Data collected from Figure 1C using PlotDigitiser. The data shows the proportion sired by the focal male (which carried either EAB or CAB genomes) after the female mated to a second male of a standard genoptype carrying marker mutations. Higher values indicate better 'sperm defence', i.e. reduced paternity loss to the second male.

#### Rice's raw data

```{r}
m <- as.data.frame(matrix(
  c(
    0.8087228560912771, 13.078892347721043,
    1.1769782848730217, 15.50243925103378,
    1.826647037173353, 23.728290665799378,
    2.1877806404122193, 19.96520001858477,
    2.8175193227824806, 19.078474190400968,
    3.191001104158999, 19.057008781303722,
    3.830622009569378, 49.462760767550996,
    4.195491350754509, 49.75561027737769),
  ncol = 2,
  byrow = TRUE)) %>% 
  mutate(treatment = rep(c("EAB", "CAB"), 4), 
         type = rep(c("all_pink", "most_pink", "most_red", "all_red"), each=2),
         n = round(V2)) %>% select(treatment, type, n)

kable(m) %>% kable_styling(full_width = F)
```

#### Effect size
To get a Cohen's d effect size, I first reduce the results from a 4x2 contingency table to a 2x2 one, by pooling the all-red and mostly-red categories, and the all-pink and mostly-pink categories. A positive effect size here would mean better sperm defence in EAB males than CAB (though the effect size does not differ significantly from zero).

```{r}
x <- matrix(c(m$n[1] + m$n[3],
         m$n[2] + m$n[4],
         m$n[5] + m$n[7],
         m$n[6] + m$n[8]), nrow = 2)

get_effect_size_2x2_matrix(x) %>% 
    mutate(Study = "Rice (1998) PNAS, 95, 6217–6221", 
           Trait = "Sperm defence", 
           Generations = "40",
           Organism = "Drosophila melanogaster",
           Interpretation = "A positive effect size means that males carrying a haplotype that underwent male-limited evolution fertilise more offspring when mating first in twice-mated females, compared to males carrying control haplotypes (though there is no significant difference between the two treatments)",
           Effect_size_source = "Chi-square test on total counts (pseudoreplicated)",
           .before = 1) %>% 
  write_csv("effect_sizes/rice98_3.csv")
```

### Sperm offence

Data collected from Figure 1D using PlotDigitiser. Rice mated females to a pink-eyed male first, then put them with a tester EAB or CAB male, and recorded the progeny genotypes qualitatively (mostly- or all-red vs pink). He reports a significant difference using a "directed $\chi^2$ contingency test" (perhaps meaning a 1-tail test), but the result appears non-significant with a standard 2-tailed test.

#### Rice's raw data

```{r}
m <- as.data.frame(matrix(
  c(
    0.8123905550128108, 10.75088339222615,
    1.1750935466627344, 16.231552691748078,
    1.8095150319809772, 2.1388484722510914,
    2.169287202079224, 0.6308459779671587,
    2.7968885366168363, 11.076699230929123,
    3.1919042966949918, 21.357825815838705,
    3.8175517501981533, 76.27520266056953,
    4.176346979779175, 61.42434005404282
  ),
  ncol = 2,
  byrow = TRUE
)) %>% 
  mutate(treatment = rep(c("EAB", "CAB"), 4), 
         type = rep(c("all_pink", "most_pink", "most_red", "all_red"), each=2),
         n = round(V2)) %>% select(treatment, type, n)

m %>% 
  pivot_wider(names_from = type, values_from = n) %>% 
  kable() %>% 
  kable_styling(full_width = F)
```


#### Effect size
To get a Cohen's d effect size, I first reduce the results from a 4x2 contingency table to a 2x2 one, by pooling the all-red and mostly-red categories, and the all-pink and mostly-pink categories. A positive effect size here would mean better sperm offence in EAB males than CAB (though the effect size does not differ significantly from zero).

```{r}
x <- matrix(c(m$n[5] + m$n[7],
         m$n[6] + m$n[8],
         m$n[1] + m$n[3],
         m$n[2] + m$n[4]), nrow = 2)

# positive effect means higher sperm offence for EAB males. It is not significant, contra Rice's analysis
get_effect_size_2x2_matrix(x) %>% 
    mutate(Study = "Rice (1998) PNAS, 95, 6217–6221", 
           Trait = "Sperm offence", 
           Generations = "40",
           Organism = "Drosophila melanogaster",
           Interpretation = "A positive effect size means that males carrying a haplotype that underwent male-limited evolution fertilise more offspring when mating second in twice-mated females, compared to males carrying control haplotypes (though there is no significant difference between the two treatments)",
           Effect_size_source = "Chi-square test on total counts (pseudoreplicated)",
           .before = 1) %>% 
  write_csv("effect_sizes/rice98_4.csv")
```

### Sex-specific development time

#### Rice's original data

These data were extracted from Figure 2A using PlotDigitizer. The data show the numbers of females eclosed in each of the 4 replicate lines over a few days. We do not know what fraction of eggs hatched, so this data tells us only the development time of females that eventually hatched. It's also unclear if more females would have hatched on later days.

```{r}
dt <- read_csv("input_data/rice98_development_time.csv", 
               show_col_types = F) %>% 
  mutate(n_females_emerged = round(prop_emerged * overall_eclosed),
         n_females_not_emerged = overall_eclosed - n_females_emerged) %>% 
  mutate(treatment = substr(population, 1, 1),
         line = population, .before = 1,
         replicate = paste(population, generation, sep = "_"), 
         day_centered = as.numeric(scale(day))) # mean-centre day and scale by dividing by its SD 

kable(dt) %>% kable_styling()
```

#### Calculating effect size

Here, I fit a binomial GLMM using the lme4 package with the form 

$$Y = day \times treatment + (1 | line)$$

With Y being a 2-column response variable giving the number of females eclosed or not yet eclosed, day in standard units, treatment being male-limited or control, and line being a 4-level factor (2 replicate lines each of the 2 treatments). I then used the fitted model to estimate the effect of treatment on the probability of eclosion (on the mean day of data collection).

The positive effect size indicates **slower** emergence of females carrying male-limited genomes (i.e. masculinisation of the phenotype). Rice interprets this result as extra female-harming mutations accumulating under male-limited selection, which is an additional valid interpretation.

```{r}
resp <- with(dt, cbind(n_females_emerged, n_females_not_emerged))
model <- glmer(
  resp ~ day_centered + treatment + (1 | line),
  family = binomial,
  data = dt,
  control = glmerControl(optimizer = "bobyqa")
)

# Get the log odds ratio effect size (+ SE) of treatment on Prob(eclosion) at the average value of day (i.e. day = 11.42)
LOR <- coef(summary(model))[2,1:2] %>% as.numeric()


tibble(
  Estimate = LOR[1] * sqrt(3) / pi,
  Var = (LOR[2]^2) * sqrt(3) / pi,
  Lower_95_CI = (LOR[1] - LOR[2]*1.96) * sqrt(3) / pi,
  Upper_95_CI = (LOR[1] + LOR[2]*1.96) * sqrt(3) / pi) %>% 
    mutate(Study = "Rice (1998) PNAS, 95, 6217–6221", 
           Trait = "Female development time", 
           Generations = "30",
           Organism = "Drosophila melanogaster",
           Interpretation = "The positive effect size means females carrying haplotypes that underwent male-limited evolution took longer to complete development than females carrying control genomes.",
           Effect_size_source = "Calculated from the raw data (digitised from a plot) using a model",
           .before = 1) %>% 
  write_csv("effect_sizes/rice98_5.csv")
```


### Female productivity

The data are from Figure 2B (captured with PlotDigitizer). Rice counted the "total number of adult offspring (measures net female fitness independent of development time) produced by EAB or CAB females (red-eyed offspring) in competition with tester-females (pink-eyed offspring)." There was no difference, implying that EAB and CAB are equally good at producing eggs and their offspring have equal survival rates (or if there are differences in these fitness components, then they cancel out). Note that in the methods and the figure axis label, Rice mentions that he only counted the daughters, while in the figure caption (quoted here) he says that all offspring were counted (meaning, sons too). It's unclear which was measured, but it likely does not qualitatively change the result.

#### Rice's original data
```{r}
x <- matrix(c(349, 344, 197, 196), nrow = 2)
colnames(x) <- c("focal_female_offspring", "rival_female_offspring")
rownames(x) <- c("CAB", "EAB")

kable(x) %>% kable_styling(full_width = F)
```

#### Calculating effect size
Here, we define a positive effect to mean that EAB females are *less* productive (which is what we'd predict if male-limited selection harms female fitness), though the observed effect size is zero or close to it.
```{r}
get_effect_size_2x2_matrix(x) %>% 
    mutate(Study = "Rice (1998) PNAS, 95, 6217–6221", 
           Trait = "Female productivity", 
           Generations = "30",
           Organism = "Drosophila melanogaster",
           Interpretation = "A positive effect size means that females carrying a haplotype that underwent male-limited evolution produced fewer offspring than females carrying a control haplotype (though there is no significant difference).",
           Effect_size_source = "Chi-square test on total counts (pseudoreplicated)",
           .before = 1) %>% 
  write_csv("effect_sizes/rice98_6.csv")
```

### Mortality of males' mates

#### Rice's original data
Data extracted from Figure 3A with PlotDigitizer. 

```{r}
mortality <- tibble(
  treatment = rep(c("EAB", "CAB"), each = 7),
  day = c(
    0,1,2,3,4,5,12, 0,1,2,3,4,5,12),
  n_dead_females = c(
    0,4,7,8,9,15,29,0,0,2,2,2,5,16),
  n_live_females = 140 - n_dead_females)

kable(mortality) %>% 
  kable_styling(full_width = F)
```


#### Calculating effect size
To calculate the effect size, I fit a Cox's proportional hazard model to the female surivivorship data, and measured the difference between the male-limited and control treatments in the per-day risk of death (conditional on being alive), taking into account the right censoring of the data (since the experiment ended on day 15, when most females were still alive). There is a weakly significant difference, such that females die faster when mated to EAB males, which is qualitatively the same conclusion Rice reached (Rice used several pairwise t-tests to test specific days one-at-a-time, which is a weaker approach, but understandable given the statistical software available in 1998). 

The positive effect size means that females die faster when mated to males carrying male-evolved compared to control genomes.

```{r}
make_surv_data <- function(day, cum_dead, treatment) {
  new_deaths <- c(cum_dead[1], diff(cum_dead))
  death_times <- rep(day, new_deaths)
  data.frame(
    time = death_times,
    event = 1,
    treatment = treatment
  )
}

surv_A <- make_surv_data(
  day = mortality$day[mortality$treatment == "EAB"],
  cum_dead = mortality$n_dead_females[mortality$treatment == "EAB"],
  treatment = "EAB"
)

surv_B <- make_surv_data(
  day = mortality$day[mortality$treatment == "CAB"],
  cum_dead = mortality$n_dead_females[mortality$treatment == "CAB"],
  treatment = "CAB"
)

surv_data <- bind_rows(surv_A, surv_B)

censored_A <- data.frame(
  time = 12,
  event = 0,
  treatment = "EAB"
)[rep(1, 140 - nrow(surv_A)), ]

censored_B <- data.frame(
  time = 12,
  event = 0,
  treatment = "CAB"
)[rep(1, 140 - nrow(surv_B)), ]

surv_data <- bind_rows(surv_data, censored_A, censored_B)

cox <- coxph(
  Surv(time, event) ~ treatment,
  data = surv_data
)

logHR <- coef(cox)["treatmentEAB"]
Var_HR <- vcov(cox)["treatmentEAB", "treatmentEAB"]
CI_logHR <- logHR + c(-1, 1) * 1.96 * sqrt(Var_HR)

d <- logHR * sqrt(3) / pi
Var_d <- Var_HR * sqrt(3) / pi
CI_d <- CI_logHR * sqrt(3) / pi

tibble(
  Estimate = d,
  Var = Var_d,
  Lower_95_CI = CI_d[1],
  Upper_95_CI = CI_d[2]) %>% 
    mutate(Study = "Rice (1998) PNAS, 95, 6217–6221", 
           Trait = "Male harmfulness to females", 
           Generations = "30",
           Organism = "Drosophila melanogaster",
           Interpretation = "A positive effect size means that males carrying a haplotype that underwent male-limited evolution have a negative effect on the survival of their mates.",
           Effect_size_source = "Calculated from the raw data (digitised from a plot) using a model",
           .before = 1) %>% 
  write_csv("effect_sizes/rice98_7.csv")
```

## Prasad et al. (2008)

Prasad, N. G., Bedhomme, S., Day, T., & Chippindale, A. K. (2007). An evolutionary cost of separate genders revealed by male-limited evolution. _The American Naturalist_, 169, 29–37.

### Male and female fitness

#### Prasad et al.'s original data

The data come from Figure 2 (transcribed with PlotDigitizer). The same size is given as n=15 per replicate population (there were 4 replicates for the control and 4 for the male-limited treatment).
```{r}
df <- data.frame(
  treatment = c("control","control", "male_evolved", "male_evolved"),
  sex = c("female", "male", "female", "male"),
  mean = c(0.858247281, 0.73376332, 0.785461824, 0.845795256),
  SE = c(0.015168311, 0.015320071, 0.015529016, 0.015670879),
  n = rep(60, 4)
)

kable(df) %>% 
  kable_styling(full_width = F)
```

#### Calculating effect size

```{r}
# For female fitness:
mes(m.1 = df$mean[df$treatment == "control" & df$sex == "female"],
    m.2 = df$mean[df$treatment == "male_evolved" & df$sex == "female"],
    sd.1 = df$SE[df$treatment == "control" & df$sex == "female"] * sqrt(60),
    sd.2 = df$SE[df$treatment == "male_evolved" & df$sex == "female"] * sqrt(60),
    n.1 = 60, n.2 = 60, verbose=F) %>%
  parse_es(study = "Prasad et al. (2007) Am. Nat. 169, 29-37",
           generations = 25,
           organism = "Drosophila melanogaster",
           trait = "Female fitness (2007)",
           interpretation = "This positive effect size indicates that females carrying male-evolved haplotypes were less fit than control females.") %>% 
  write_csv("effect_sizes/prasad2007_1.csv")

# For male fitness:
mes(m.2 = df$mean[df$treatment == "control" & df$sex == "male"],
    m.1 = df$mean[df$treatment == "male_evolved" & df$sex == "male"],
    sd.2 = df$SE[df$treatment == "control" & df$sex == "male"] * sqrt(60),
    sd.1 = df$SE[df$treatment == "male_evolved" & df$sex == "male"] * sqrt(60),
    n.1 = 60, n.2 = 60, verbose=F) %>%
  parse_es(study = "Prasad et al. (2007) Am. Nat. 169, 29-37",
           generations = 25,
           organism = "Drosophila melanogaster",
           trait = "Male fitness (2007)",
           interpretation = "This positive effect size indicates that males carrying male-evolved haplotypes were fitter than control males.") %>% 
  write_csv("effect_sizes/prasad2007_2.csv")
```

### Development time

#### Prasad et al.'s original data

These data (development time in hours) are presented in prose in the Results. The sample size is not recorded, so we estimated it as n=60 flies per treatment (a feasible number of individuals on the same order as the same study's fitness assays). An error here would over- or under-inflate the effect size precision, but likely not have a major effect on the overall effect size estimate.

```{r}
df <- data.frame(
  treatment = c("control","control", "male_evolved", "male_evolved"),
  sex = c("female", "male", "female", "male"),
  mean = c(224.8, 225.9, 227.0, 229.8),
  SE = c(0.61, 0.59, 0.57, 0.54),
  n = rep(60, 4)
)

kable(df) %>% 
  kable_styling(full_width = F)
```

#### Calculating effect size

```{r}
# For female DT:
mes(m.2 = df$mean[df$treatment == "control" & df$sex == "female"],
    m.1 = df$mean[df$treatment == "male_evolved" & df$sex == "female"],
    sd.2 = df$SE[df$treatment == "control" & df$sex == "female"] * sqrt(60),
    sd.1 = df$SE[df$treatment == "male_evolved" & df$sex == "female"] * sqrt(60),
    n.1 = 60, n.2 = 60, verbose=F) %>%
  parse_es(study = "Prasad et al. (2007) Am. Nat. 169, 29-37",
           generations = 25,
           organism = "Drosophila melanogaster",
           trait = "Female development time",
           interpretation = "This positive effect size indicates that females carrying male-evolved haplotypes developed faster than control females  (i.e. have a masculinised phenotype).") %>% 
  write_csv("effect_sizes/prasad2007_3.csv")

# For male DT:
mes(m.2 = df$mean[df$treatment == "control" & df$sex == "male"],
    m.1 = df$mean[df$treatment == "male_evolved" & df$sex == "male"],
    sd.2 = df$SE[df$treatment == "control" & df$sex == "male"] * sqrt(60),
    sd.1 = df$SE[df$treatment == "male_evolved" & df$sex == "male"] * sqrt(60),
    n.1 = 60, n.2 = 60, verbose=F) %>%
  parse_es(study = "Prasad et al. (2007) Am. Nat. 169, 29-37",
           generations = 25,
           organism = "Drosophila melanogaster",
           trait = "Male development time",
           interpretation = "This positive effect size indicates that males carrying male-evolved haplotypes developed faster than control males  (i.e. have a masculinised phenotype).") %>% 
  write_csv("effect_sizes/prasad2007_4.csv")
```

### Dry mass

#### Prasad et al.'s original data

These data (dry mass in micrograms) are presented in prose in the Results. The sample size is not recorded, so we estimated it as n=60 flies per treatment (a feasible number of individuals on the same order as the same study's fitness assays). An error here would over- or under-inflate the effect size precision, but likely not have a major effect on the overall effect size estimate.

```{r}
df <- data.frame(
  treatment = c("control","control", "male_evolved", "male_evolved"),
  sex = c("female", "male", "female", "male"),
  mean = c(275.9, 217.9, 267.4, 212.5),
  SE = c(1.45, 1.25, 1.82, 1.43),
  n = rep(60, 4)
)

kable(df) %>% 
  kable_styling(full_width = F)
```

#### Calculating effect size

```{r}
# For female dry mass:
mes(m.1 = df$mean[df$treatment == "control" & df$sex == "female"],
    m.2 = df$mean[df$treatment == "male_evolved" & df$sex == "female"],
    sd.1 = df$SE[df$treatment == "control" & df$sex == "female"] * sqrt(60),
    sd.2 = df$SE[df$treatment == "male_evolved" & df$sex == "female"] * sqrt(60),
    n.1 = 60, n.2 = 60, verbose=F) %>%
  parse_es(study = "Prasad et al. (2007) Am. Nat. 169, 29-37",
           generations = 25,
           organism = "Drosophila melanogaster",
           trait = "Female body mass (2007)",
           interpretation = "This positive effect size indicates that females carrying male-evolved haplotypes are lighter than control females (i.e. have a masculinised phenotype).") %>% 
  write_csv("effect_sizes/prasad2007_5.csv")

# For male dry mass:
mes(m.1 = df$mean[df$treatment == "control" & df$sex == "male"],
    m.2 = df$mean[df$treatment == "male_evolved" & df$sex == "male"],
    sd.1 = df$SE[df$treatment == "control" & df$sex == "male"] * sqrt(60),
    sd.2 = df$SE[df$treatment == "male_evolved" & df$sex == "male"] * sqrt(60),
    n.1 = 60, n.2 = 60, verbose=F) %>%
  parse_es(study = "Prasad et al. (2007) Am. Nat. 169, 29-37",
           generations = 25,
           organism = "Drosophila melanogaster",
           trait = "Male body mass (2007)",
           interpretation = "This positive effect size indicates that males carrying male-evolved haplotypes are lighter than control males (i.e. have a masculinised phenotype).") %>% 
  write_csv("effect_sizes/prasad2007_6.csv")
```





## Bedhomme et al. (2008)

Bedhomme, S., Prasad, N. G., Jiang, P. P., & Chippindale, A. K. (2008). Reproductive behaviour evolves rapidly when intralocus sexual conflict is removed. _PLOS ONE_, 3, e2187.

Note: these authors also recorded the number of matings observed (in mixed-sex vials where either the males or the females carried either male-evolved or control genomes, while the other sex was a standard genotype). We did not include these data, since the authors mention that the observed number of matings probably differs substantially from the actual number of matings. They also measured the number of flies present near the yeast food source in those vials, but they did not score the sex of the flies at the yeast - this makes it unclear whether the two genome types directly affecting feeding rate, or something else (e.g. they might affect social behaviours in a way that causes space use near the food). These yeast data are thus difficult to interpret and we leave them out of the meta-analysis.

This leaves the data on courthsip. The authors measured the number of observed courtship events in A) vials where the males carried either of the 2 evolved genomes and females were a standard genotype, and B) vials where the females carried either of the 2 evolved genomes and males were a standard genotype. For brevity, we report these two traits as 'male courtship rate' and 'female attractivenss' respectively, though these names are a little simplistic and hide some of the nuance in interpreting these data.

### Male courtship rate

#### Bedhomme et al.'s original data

The data come from Table 1. The means represent the number of courtship events towards females per vial, in vials where the males carried either MLE or control genomes (females were a standard genotype). There were 40 vials in total, 10 vials per evolving population, with 4 replicate populations of the MLE treatment and 4 of the control.

Note that the authors were surprised to see LESS courtship by the MLE males, though they note they mate just as often as controls, and rationalised _post hoc_ that the lower courtship reflects less need to engage in courtship in order to mate. We therefor have coded this effect size as positive even though it should arguably be negative, since it went against the _a priori_ expectation under IaSC. Changing this effect to negative does not change the overall positive, significant effect size in the meta-analysis (it just makes it slightly less positive, but still confidently above zero).

```{r}
# The authors were surprised to see LESS courtship by the MLE males, though they note they mate just as often as controls,
# and suggest this finding reflects increased attractiveness of MLE males so they do not need to court as much

df <- tibble(
  mean = c(13.6, 22.03),
  SE = c(1.13, 1.14),
  SD = SE * sqrt(40),
  n = c(40,40),
  treatment = c("MLE", "control")
)
```

#### Calculating effect size
```{r}
mes(m.2 = df$mean[1], m.1 = df$mean[2],
    sd.2 = df$SD[1], sd.1 = df$SD[2],
    n.1 = 40, n.2 = 40, verbose = F) %>% 
  parse_es(study = "Bedhomme et al. (2008) PLOS ONE 3, e2187",
           generations = 52,
           organism = "Drosophila melanogaster",
           trait = "Male courtship rate",
           interpretation = "This negative effect size indicates that males carrying male-evolved haplotypes courted females less often than males carrying control haplotypes. This might be because the former males are more attractive and thus need to court less."
          ) %>% 
  write_csv("effect_sizes/bedhomme2008_1.csv")
```

### Female attractiveness

#### Bedhomme et al.'s original data

The data come from Table 1. The means represent the number of courtship events towards females per vial, in vials where the females carried either MLE or control genomes (males were a standard genotype). There were 40 vials in total, 10 vials per evolving population, with 4 replicate populations of the MLE treatment and 4 of the control.

```{r}
# The MLE females are courted slightly less, implying they are less attractive in some way (pheromones, behaviour, smaller body size?)

df <- tibble(
  mean = c(4.65, 8),
  SE = c(0.66, 1.31),
  SD = SE * sqrt(40),
  n = c(40,40),
  treatment = c("MLE", "control")
)
```

#### Calculating effect size
```{r}
mes(m.2 = df$mean[1], m.1 = df$mean[2],
    sd.2 = df$SD[1], sd.1 = df$SD[2],
    n.1 = 40, n.2 = 40, verbose = F) %>% 
  parse_es(study = "Bedhomme et al. (2008) PLOS ONE 3, e2187",
           generations = 52,
           organism = "Drosophila melanogaster",
           trait = "Female attractiveness",
           interpretation = "This positive effect size indicates that females carrying male-evolved haplotypes were courted less often by standard males (perhaps due to a masculinisation of the phenotype after male-limited evolution?), compared to females carrying control haplotypes.") %>% 
  write_csv("effect_sizes/bedhomme2008_2.csv")
```


## Abbott et al. (2010) 

Abbott, J. K., Bedhomme, S., & Chippindale, A. K. (2010). Sexual conflict in wing size and shape in _Drosophila melanogaster_. _Journal of Evolutionary Biology_, 23(9), 1989–1997.

The means for each sex/trait/treatment combination come from the Dryad supplementary material. That supplement gives the sample sizes and means for each of the 8 replicate evolution lines (not raw data), so I took a weighted average (weighting by sample size) across the 4 lines within each treatment to get an overall mean for both treatments, for three male and three female traits (body mass, wing size, and fitness). I also got the sample sizes from this same file.

To get the standard deviations needed for calculating effect size, I used PlotDigitizer to get the standard error of the mean from the error bars in Figure 2, then multiplied these by $\sqrt{N}$. I then calculated Cohen's d from the mean, SD and N using the function `compute.es::mes()`. Positive effect sizes mean the male-limited evolution caused phenotypic masculinisation (i.e. _smaller_ body mass and wing size, _higher_ male fitness, and _lower_ female fitness).


### Abbott et al.'s original data

```{r}
abb <- read_excel("input_data/Abbottetal.JEBdryaddata.xlsx") %>% 
  select(treatment = `Selection treatment`,
         sex = Sex, N, body_mass = Mass, wing_size = `Wing Size`, fitness = `Relative fitness`) %>% 
  filter(!is.na(sex)) %>% 
  group_by(sex, treatment) %>% 
  summarise(
    body_mass = sum(body_mass * N) / sum(N),
    wing_size = sum(wing_size * N) / sum(N),
    fitness = sum(fitness * N) / sum(N),
    N = sum(N), .groups = "drop") %>% 
  mutate(bodymass_SE = c(0.42645353268463704 - 0.4156004674677981, # digitised from error bars in Figure S1A
                     0.3828284977515815 - 0.3764849724346434,
                     0.24649113335535175 - 0.24378064581692538,
                     0.2303028378344046 - 0.22852442773303525)) %>% 
  mutate(wingsize_SE = c(2.328553386369538 - 2.3147954486437996, # digitised from error bars in Figure 2a
                     2.3005653785321405 - 2.2892625292131914,
                     2.0457272019074146 - 2.036586718915984,
                     2.021764122659994 - 2.008893793819787)) %>% 
  mutate(fitness_SE = c(1.074087075863868 - 1.0575940429537967, # digitised from error bars in Figure 2c
                     0.9990512054171464 - 0.9711295362645963,
                     1.0687912935478567 - 1.0388472684583931,
                     0.9872709537285416 - 0.9570842005569136)) %>% 
  mutate(bodymass_SD = bodymass_SE * sqrt(N),
         wingsize_SD = wingsize_SE * sqrt(N),
         fitness_SD = fitness_SE * sqrt(N))
  
abb %>% 
  kable() %>% 
  kable_styling(full_width = F)
```

### Female body mass effect size

```{r}
mes(m.1 = abb$body_mass[1], m.2 = abb$body_mass[2],
    sd.1 = abb$bodymass_SD[1], sd.2 = abb$bodymass_SD[2],
    n.1 = abb$N[1], n.2 = abb$N[2], verbose=F) %>% 
  parse_es(study = "Abbott et al. (2010) J. Evol. Biol. 23, 1989–1997", 
           trait = "Female body mass (2010)", 
           generations = 82, 
           organism = "Drosophila melanogaster", 
           interpretation = "The positive effect size indicates that females carrying male-evolved haplotypes have lower body mass than females carrying control genotypes.") %>% 
  write_csv("effect_sizes/abbott2010_1.csv")
```

### Male body mass effect size

```{r}
mes(m.1 = abb$body_mass[3], m.2 = abb$body_mass[4],
    sd.1 = abb$bodymass_SD[3], sd.2 = abb$bodymass_SD[4],
    n.1 = abb$N[3], n.2 = abb$N[4], verbose=F) %>% 
  parse_es(study = "Abbott et al. (2010) J. Evol. Biol. 23, 1989–1997", 
           trait = "Male body mass (2010)", 
           generations = 82, 
           organism = "Drosophila melanogaster", 
           interpretation = "The positive effect size indicates that males carrying male-evolved haplotypes have lower body mass than males carrying control genotypes.") %>% 
  write_csv("effect_sizes/abbott2010_2.csv")
```

### Female wing size effect size

```{r}
mes(m.1 = abb$wing_size[1], m.2 = abb$wing_size[2],
    sd.1 = abb$wingsize_SD[1], sd.2 = abb$wingsize_SD[2],
    n.1 = abb$N[1], n.2 = abb$N[2], verbose=F) %>% 
  parse_es(study = "Abbott et al. (2010) J. Evol. Biol. 23, 1989–1997", 
           trait = "Female wing size", 
           generations = 82, 
           organism = "Drosophila melanogaster", 
           interpretation = "A positive effect size would indicate that females carrying male-evolved haplotypes have smaller wings than females carrying control genotypes (though wing size does not differ consistently).") %>% 
  write_csv("effect_sizes/abbott2010_3.csv")
```

### Male wing size effect size

```{r}
mes(m.1 = abb$wing_size[3], m.2 = abb$wing_size[4],
    sd.1 = abb$wingsize_SD[3], sd.2 = abb$wingsize_SD[4],
    n.1 = abb$N[3], n.2 = abb$N[4], verbose=F) %>% 
  parse_es(study = "Abbott et al. (2010) J. Evol. Biol. 23, 1989–1997", 
           trait = "Male wing size", 
           generations = 82, 
           organism = "Drosophila melanogaster", 
           interpretation = "A positive effect size would indicate that males carrying male-evolved haplotypes have smaller wings than males carrying control genotypes (though wing size does not differ consistently).") %>% 
  write_csv("effect_sizes/abbott2010_4.csv")
```

### Female fitness effect size

```{r}
mes(m.1 = abb$fitness[1], m.2 = abb$fitness[2],
    sd.1 = abb$fitness_SD[1], sd.2 = abb$fitness_SD[2],
    n.1 = abb$N[1], n.2 = abb$N[2], verbose=F) %>% 
  parse_es(study = "Abbott et al. (2010) J. Evol. Biol. 23, 1989–1997", 
           trait = "Female fitness (2010)", 
           generations = 82, 
           organism = "Drosophila melanogaster", 
           interpretation = "The positive effect size indicates that females carrying male-evolved haplotypes have lower fitness thanthan females carrying control genotypes.") %>% 
  write_csv("effect_sizes/abbott2010_5.csv")
```

### Male fitness effect size

```{r}
mes(m.2 = abb$fitness[3], m.1 = abb$fitness[4],
    sd.2 = abb$fitness_SD[3], sd.1 = abb$fitness_SD[4],
    n.2 = abb$N[3], n.1 = abb$N[4], verbose=F) %>% 
  parse_es(study = "Abbott et al. (2010) J. Evol. Biol. 23, 1989–1997", 
           trait = "Male fitness (2010)", 
           generations = 82, 
           organism = "Drosophila melanogaster", 
           interpretation = "The positive effect size indicates that males carrying male-evolved haplotypes have higher fitness than males carrying control genotypes.") %>% 
  write_csv("effect_sizes/abbott2010_6.csv")
```



## Jiang et al. (2011)

Jiang, P. P., Bedhomme, S., Prasad, N. G., & Chippindale, A. K. (2011). Sperm competition and mate harm unresponsive to male-limited selection in _Drosophila_: An evolving genetic architecture under domestication. _Evolution_, 65(9), 2448–2460.

### Mating defence

#### Jiang et al.'s original data

These data were transcribed from the right 2 columns in Figure 2A (i.e. using the data for LH_ST females, since these better measure the effects of male-limitation rather than adaptation to the clone generator females) using PlotDigitiser. The sample size is estimated based on the overall number of replicates being 12.6 (mentioned in the text) with 10 females per replicate. This is a bit pseudo-replicated, overstating the precision of the effect size. 

```{r}
df <- tibble(
  mean = c(0.6349673427294603, 0.5952973530422826),
  SE = c(0.6734272946029564 - 0.6349673427294603, 
         0.628030250945342 - 0.5952973530422826),
  SD = SE * sqrt(126),
  N = c(126, 126),
  treatment = c("MLE", "control")
)

df %>% 
  kable() %>% 
  kable_styling(full_width = F)
```

### Calculating effect size

```{r}
mes(m.1 = df$mean[2], m.2 = df$mean[1],
    sd.1 = df$SD[2], sd.2 = df$SD[1],
    n.1 = df$N[2], n.2 = df$N[1], verbose = F) %>% 
  parse_es(study = "Jiang et al. (2011) Evolution 65: 2448-2460", 
           trait = "Mating defence", 
           generations = 65, 
           organism = "Drosophila melanogaster", 
           interpretation = "A positive effect size would mean that the mates of males carrying haplotypes that had undergone male-limited evolution are less likely to remate with a new male, compared to the mates of control males (the effect goes non-significantly in the opposite direction, though).") %>% 
   write_csv("effect_sizes/jiang_2011_1.csv")
```

### Sperm defence

#### Jiang et al.'s original data

These data were transcribed from the right 2 columns in Figure 2B (i.e. using the data for LH_ST females, since these better measure the effects of male-limitation rather than adaptation to the clone generator females) using PlotDigitiser. Same sample size estimate as before.

```{r}
df <- tibble(
  mean = c(0.293426235316184, 0.2820820973803586),
  SE = c(0.30766024550020493 - 0.293426235316184, 
         0.29350959701565116 - 0.2820820973803586),
  SD = SE * sqrt(126),
  N = c(126, 126),
  treatment = c("MLE", "control")
)

df %>% 
  kable() %>% 
  kable_styling(full_width = F)
```

### Calculating effect size

```{r}
mes(m.1 = df$mean[1], m.2 = df$mean[2],
    sd.1 = df$SD[1], sd.2 = df$SD[2],
    n.1 = df$N[1], n.2 = df$N[2], verbose = F) %>% 
  parse_es(study = "Jiang et al. (2011) Evolution 65: 2448-2460", 
           trait = "Sperm defence", 
           generations = 65, 
           organism = "Drosophila melanogaster", 
           interpretation = "A positive effect size would mean that males carrying haplotypes that had undergone male-limited evolution sire more offspring when mating first with twice-mated females, compared to control males (the effect goes non-significantly in the opposite direction, though).") %>% 
   write_csv("effect_sizes/jiang_2011_2.csv")
```

### Male mating success

#### Jiang et al.'s original data

These data were transcribed from the right 2 columns in Figure 3A (i.e. using the data for LH_ST females, since these better measure the effects of male-limitation rather than adaptation to the clone generator females) using PlotDigitiser. Same sample size estimate as before. Note this trait is called 'Mating offence' in the paper since the females had previously mated, but we've renamed it here for clarity outside the context of Jiang et al. and consistency with other traits in the meta-analysis.

```{r}
df <- tibble(
  mean = c(0.7297174519489161, 0.730766769892253),
  SE = c(0.7630716863307143 - 0.7297174519489161, 
         0.7582729762239909 - 0.730766769892253),
  SD = SE * sqrt(126),
  N = c(126, 126),
  treatment = c("MLE", "control")
)

df %>% 
  kable() %>% 
  kable_styling(full_width = F)
```

### Calculating effect size

```{r}
mes(m.1 = df$mean[1], m.2 = df$mean[2],
    sd.1 = df$SD[1], sd.2 = df$SD[2],
    n.1 = df$N[1], n.2 = df$N[2], verbose = F) %>% 
  parse_es(study = "Jiang et al. (2011) Evolution 65: 2448-2460", 
           trait = "Male mating success", 
           generations = 65, 
           organism = "Drosophila melanogaster", 
           interpretation = "A positive effect size would mean that the males carrying haplotypes that had undergone male-limited evolution are more likely to mate with females, compared to the mates of control males (there is no difference in mating rates, though).") %>% 
   write_csv("effect_sizes/jiang_2011_3.csv")
```

### Sperm offence

#### Jiang et al.'s original data

These data were transcribed from the right 2 columns in Figure 3B (i.e. using the data for LH_ST females, since these better measure the effects of male-limitation rather than adaptation to the clone generator females) using PlotDigitiser. Same sample size estimate as before. 

```{r}
df <- tibble(
  mean = c(0.8227287156371025, 0.8270903434702559),
  SE = c(0.8284807313821345 - 0.8227287156371025, 
         0.8337502380801219 - 0.8270903434702559),
  SD = SE * sqrt(126),
  N = c(126, 126),
  treatment = c("MLE", "control")
)

df %>% 
  kable() %>% 
  kable_styling(full_width = F)
```

### Calculating effect size

```{r}
mes(m.1 = df$mean[1], m.2 = df$mean[2],
    sd.1 = df$SD[1], sd.2 = df$SD[2],
    n.1 = df$N[1], n.2 = df$N[2], verbose = F) %>% 
  parse_es(study = "Jiang et al. (2011) Evolution 65: 2448-2460", 
           trait = "Sperm offence", 
           generations = 65, 
           organism = "Drosophila melanogaster", 
           interpretation = "Apositive effect size would mean that males carrying haplotypes that had undergone male-limited evolution sire more offspring when mating second with twice-mated females, compared to control males (there is no difference in mating rates, though).") %>% 
   write_csv("effect_sizes/jiang_2011_4.csv")
```


## Bedhomme et al. (2011)

Bedhomme, S., Chippindale, A. K., Prasad, N. G., Delcourt, M., Abbott, J. K., Mallet, M. A., & Rundle, H. D. (2011). Male-limited evolution suggests no extant intralocus sexual conflict over the sexually dimorphic cuticular hydrocarbons of _Drosophila melanogaster_. _Journal of Genetics_, 90(3), 443–452.

This paper measured the abundance of dozens of hydrocarbons present on the body surface of both sexes, in the male-limited evolution treatment or the control lines. The authors analysed each hydrocarbon with its own univariate test, and largely found no differences between treatments. 

Since there are many hydrocarbons, we have not included them all here: we focus on a single male-unique hydrocarbon, 7-C$_{23:1}$, which is known to function as a male sex pheromone (e.g. Grillet et al., 2006; https://doi.org/10.1098/rspb.2005.3332), because this seems the most likely one that might evolve following removal of selection on females. The means and standard errors in the code chunk below come from row M7 of Table 1, and the sample sizes are inferred from the methods (30 males per selection line, 4 selection lines per treatment = 120 males per treatment assuming no missing samples). The positive (non-significant) effect size implies more sex pheromones in the ML males.

The authors also carried out multivariate tests on the first 3 principal components, but the results for those tests are not detailed enough to include in this meta-analysis. The multivariate tests suggest no difference in CHCs between treatments, for either males or females.

### Bedhomme et al.'s original data 

```{r}
tibble(treatment = c("Control", "Male-limited evolution"),
       mean = c(34.518, 35.003),
       SE = c(0.351, 0.322),
       N = c(120, 120)) %>% 
  kable() %>% kable_styling(full_width = F)
```

### Calculating effect size

```{r}
mes(m.2 = 34.518, m.1 = 35.003,
    sd.2 = 0.351 * sqrt(120), sd.1 = 0.322 * sqrt(120),
    n.1 = 120, n.2 = 120, verbose = F) %>% 
  parse_es(study = "Bedhomme et al. (2011) J. Genetics 90, 443–452", 
           trait = "Male 7-C23:1 sex pheromone", 
           generations = 82, 
           organism = "Drosophila melanogaster", 
           interpretation = "A positive effect size would imply that males carrying male-limited haplotypes produce more 7-C23:1 than males carrying control genomes (though there is no significant difference).") %>% 
  write_csv("effect_sizes/bedhomme2011.csv")
```


## Abbott et al. (2013)

Abbott, J. K., Innocenti, P., Chippindale, A. K., & Morrow, E. H. (2013). Epigenetics and sex-specific fitness: An experimental test using male-limited evolution in _Drosophila melanogaster_. _PLOS ONE_, 8, e70493.

Here we focus on the effect size comparing the CDX and MLX treatments, since CDX is a stronger control than the other group (labelled C).

Two traits are recorded here: 1) male fitness, and 2) the proportion of males among the progeny that survived to adulthood. Both have positive effect sizes, indicating higher values in the male-limited (MLX) treatment relative to the control (CDX). 

For the second trait, the excess of males among surviving progeny implies either a negative genetic effect of male-limited evolution on female survival to adulthood, or some sort of non-genetic effect of the father's genotype on offspring sex ratio - it cannot be that the male-limited X has a positive genetic effect on male survival (since males do not pass on the X chromosome to their sons).

### Abbott et al.'s original data

The data were extracted from Figures 3 and S4 using PlotDigitizer. I have assumed that the standard errors were calculated using individual vials (rather that experimental evolution lines) as replicates, given the small error bars, for a sample of 60 per treatment (20 for each of 3 replicate lines).

```{r}
df <- tibble(
  trait = c("Male fitness", "Male fitness", "Proportion of sons", "Proportion of sons"),
  treatment = c("CDX", "MLX", "CDX", "MLX"),
  mean  = c(0.7781242592641365, 1.2066580631567856, 0.3816372189736579, 0.4524972141486783),
  SE = c(0.8454581369011561 - 0.7781242592641365,
         1.2642102768047616 - 1.2066580631567856,
         0.38419101175227177 - 0.3816372189736579, 
         0.45990786586556687 - 0.4524972141486783),
  N = c(60, 60, 60, 60)) 

df %>% 
  kable() %>% 
  kable_styling(full_width = F)
```


### Male fitness

Fitness was measured as the proportion of adult offspring sired when in competition for matings with marked competitor males.

```{r}
df <- df %>% mutate(SD = SE * sqrt(N))

mes(m.1 = df$mean[2], m.2 = df$mean[1],
    sd.1 = df$SD[2], sd.2 = df$SD[1],
    n.1 = df$N[2], n.2 = df$N[1], verbose = F) %>% 
  parse_es(study = "Abbott et al. (2013) PLOS ONE 8, e70493", 
           trait = "Male fitness (2013)", 
           generations = 40, 
           organism = "Drosophila melanogaster", 
           interpretation = "Male fitness is higher in males carrying an X chromosome that underwent male-limited evolution compared to males carrying an X that experienced only 1 generation of father-to-son inheritance.") %>% 
  write_csv("effect_sizes/abbott_2013_1.csv")
```


```{r}
mes(m.1 = df$mean[4], m.2 = df$mean[3],
    sd.1 = df$SD[4], sd.2 = df$SD[3],
    n.1 = df$N[4], n.2 = df$N[3], verbose = F) %>% 
  parse_es(study = "Abbott et al. (2013) PLOS ONE 8, e70493", 
           trait = "Male effect on % sons (2013)", 
           generations = 40, 
           organism = "Drosophila melanogaster", 
           interpretation = "The proportion of sons among the progeny of males carrying an X chromosome that underwent male-limited evolution is higher than that of males carrying an X that experienced only 1 generation of father-to-son inheritance.") %>% 
  write_csv("effect_sizes/abbott_2013_2.csv")
```


## Abbott et al. (2020)

Abbott, J. K., Chippindale, A. K., & Morrow, E. H. (2020). The microevolutionary response to male-limited X-chromosome evolution in _Drosophila melanogaster_ reflects macroevolutionary patterns. _Journal of Evolutionary Biology_, 33, 738–750.

All data transcribed using PlotDigitiser from Figures 2A, 2B, S1 and S2. Sample sizes given in the Methods, bottom left of page 741.

When deciding on coding the sign of the effect sizes, I considered high locomotion a male-typical trait and so predicted both female and male locomotion to increase in the MLX treatment. I also predicted increased offspring survival in MLX, due to the unmasking of deleterious alleles in males, and an elevated proportion of sons (whether the X was present in the mother or the father) due to the removal of selection on female-harming alleles and (in the case of the X coming from the mother) extra purging of male-harming alleles.

### Abbott et al.'s original data

```{r}
df <- tibble(
  trait = c("Female fitness", "Female fitness",
            "Male fitness", "Male fitness", 
            "Female locomotion", "Female locomotion",
            "Male locomotion", "Male locomotion",
            "Female effect on % sons", "Female effect on % sons",
            "Male effect on % sons", "Male effect on % sons",
            "Egg-to-adult survival", "Egg-to-adult survival"),
  treatment = rep(c("CDX", "MLX"), 7),
  mean  = c(0.9502716648321987, 1.0522634062212999, 
            0.8833110159213118, 1.117530268720336,
            0.13367807311356172, 0.33332353958957683,
            0.44689179218981584, 0.3798242349453836,
            0.18625256770849793, 0.34924785892614485,
            0.3340217425654963, 0.4002676737351073,
            0.4647589355305646,0.5514258496331762),
  SE = c(1.0482869341484446 - 0.9502716648321987,
         1.0971487461353486 - 1.0522634062212999,
         0.9376523952074818 - 0.8833110159213118, 
         1.1710079291077475 - 1.117530268720336,
         0.16792026586749717 - 0.13367807311356172,
         0.367418826187165 - 0.33332353958957683,
         0.4805985936183966 - 0.44689179218981584,
         0.3798242349453836 - 0.3454253422913443,
         0.20611478052017823 - 0.18625256770849793,
         0.3695196409948488 - 0.34924785892614485,
         0.34469898555762735 - 0.3340217425654963,
         0.4139503207660462 - 0.4002676737351073,
         0.5538655680510708 - 0.4647589355305646,
         0.6240087097452648 - 0.5514258496331762),
  N = c(rep(60, 4), rep(30, 4), rep(60, 6))) 

df %>% 
  kable() %>% kable_styling(full_width = F)
```

### Female fitness effect size

```{r}
df <- df %>% mutate(SD = SE * sqrt(N))

mes(m.1 = df$mean[1], m.2 = df$mean[2],
    sd.1 = df$SD[1], sd.2 = df$SD[2],
    n.1 = df$N[1], n.2 = df$N[2], verbose = F) %>% 
  parse_es(study = "Abbott et al. (2020) J. Evol. Biol. 33, 738–750", 
           trait = "Female fitness", 
           generations = 40, 
           organism = "Drosophila melanogaster", 
           interpretation = "A positive effect size would mean that females carrying a male-evolved X chromosome have lower fitness than females carrying a control genome (the effect is non-significantly negative).") %>% 
  write_csv("effect_sizes/abbott_2020_1.csv")
```

### Male fitness effect size

```{r}
mes(m.1 = df$mean[4], m.2 = df$mean[3],
    sd.1 = df$SD[4], sd.2 = df$SD[3],
    n.1 = df$N[4], n.2 = df$N[3], verbose = F) %>% 
  parse_es(study = "Abbott et al. (2020) J. Evol. Biol. 33, 738–750", 
           trait = "Male fitness (2020)", 
           generations = 40, 
           organism = "Drosophila melanogaster", 
           interpretation = "The positive effect size indicates that males carrying a male-evolved X chromosome have higher fitness than males carrying a control genome.") %>% 
  write_csv("effect_sizes/abbott_2020_2.csv")
```

### Female locomotion

```{r}
mes(m.1 = df$mean[6], m.2 = df$mean[5],
    sd.1 = df$SD[6], sd.2 = df$SD[5],
    n.1 = df$N[6], n.2 = df$N[5], verbose = F) %>% 
  parse_es(study = "Abbott et al. (2020) J. Evol. Biol. 33, 738–750", 
           trait = "Female locomotion", 
           generations = 40, 
           organism = "Drosophila melanogaster", 
           interpretation = "The positive effect size indicates that females carrying a male-evolved X chromosome have greater locomotory activity than females with a control genome.") %>% 
  write_csv("effect_sizes/abbott_2020_3.csv")
```

### Male locomotion

```{r}
mes(m.1 = df$mean[8], m.2 = df$mean[7],
    sd.1 = df$SD[8], sd.2 = df$SD[7],
    n.1 = df$N[8], n.2 = df$N[7], verbose = F) %>% 
  parse_es(study = "Abbott et al. (2020) J. Evol. Biol. 33, 738–750", 
           trait = "Male locomotion", 
           generations = 40, 
           organism = "Drosophila melanogaster", 
           interpretation = "A positive effect size would indicate that males carrying a male-evolved X chromosome have greater locomotory activity than males with a control genome (there is a non-significant trend in the opposite direction, though).") %>% 
  write_csv("effect_sizes/abbott_2020_4.csv")
```

### Female effect on % sons

```{r}
mes(m.1 = df$mean[10], m.2 = df$mean[9],
    sd.1 = df$SD[10], sd.2 = df$SD[9],
    n.1 = df$N[10], n.2 = df$N[9], verbose = F) %>% 
  parse_es(study = "Abbott et al. (2020) J. Evol. Biol. 33, 738–750", 
           trait = "Female effect on % sons  (2020)", 
           generations = 40, 
           organism = "Drosophila melanogaster", 
           interpretation = "The positive effect size indicates that females carrying a male-evolved X chromosome have a greater proportion of sons among their progeny that reach adulthood (indicating enhanced male survival, reduced female survival, and/or changes to the primary sex ratio). Note that all sons and 50% daughters actually inherit the male-evolved X from the tester females.") %>% 
  write_csv("effect_sizes/abbott_2020_5.csv")
```

### Male effect on % sons

```{r}
mes(m.1 = df$mean[12], m.2 = df$mean[11],
    sd.1 = df$SD[12], sd.2 = df$SD[11],
    n.1 = df$N[12], n.2 = df$N[11], verbose = F) %>% 
  parse_es(study = "Abbott et al. (2020) J. Evol. Biol. 33, 738–750", 
           trait = "Male effect on % sons", 
           generations = 40, 
           organism = "Drosophila melanogaster", 
           interpretation = "The positive effect size indicates that males carrying a male-evolved X chromosome have a greater proportion of sons among their progeny that reach adulthood (indicating enhanced male survival, reduced female survival, and/or changes to the primary sex ratio). Note that all daughters and none of the sons actually inherit the male-evolved X from the tester males.") %>% 
  write_csv("effect_sizes/abbott_2020_6.csv")
```

### Egg-to-adult survival

```{r}
mes(m.1 = df$mean[14], m.2 = df$mean[13],
    sd.1 = df$SD[14], sd.2 = df$SD[13],
    n.1 = df$N[14], n.2 = df$N[13], verbose = F) %>% 
  parse_es(study = "Abbott et al. (2020) J. Evol. Biol. 33, 738–750", 
           trait = "Egg-to-adult survival", 
           generations = 40, 
           organism = "Drosophila melanogaster", 
           interpretation = "A positive effect size would indicates elevated egg-to-adult survival in the progeny of females carrying a male-evolved X chromosome relative to females carrying control genomes. Note that all sons and 50% daughters actually inherit the male-evolved X from the tester females. There is seemingly no difference based on the available data, though the authors find that MLX has greater survival than the control in a quasibinomial GLM presented in the supplement.") %>% 
  write_csv("effect_sizes/abbott_2020_5.csv")
```


## Lund-Hansen et al. (2020)

Lund-Hansen, K. K., Abbott, J. K., & Morrow, E. H. (2020). Feminization of complex traits in _Drosophila melanogaster_ via female-limited X chromosome evolution. _Evolution_, 74, 2703–2713.

Here, I focus on comparisons between the female-limited X chromosome treatment and the 'control FM' treatment (ignoring the wild-type control), since differences between the former 2 treatments are expected to arise due to female-limited inheritance of the X (rather than adaption to the FM balancer chromosome). I also focus on the later measurements made around generations 39-41, rather than the early measurements, since the late measurements have given more time for experimental evolution to produce evolution.

Effect sizes are calculated from models that motly follow those in the authors' archived code, with a couple of differences. First, I did not run one model for both sexes (with sex as a predictor) as in the original, but rather split the data by sex and ran two models (running with both sexes together gave warnings about rank deficiency). Second, I treated the experimental replicate lines as fixed rather than random factors (since there are only 3 replicates per treatment, and it's recommended to have more levels in order to use random factors). The results are very similar to those in the paper. Third, for the development time data, the response variable (time until development) was not normal, as is common for time-to-event variables, and so I used a Cox PH model instead of linear (mixed) model as in the original paper. The treatment effect (as a proportional hazard ratio) was converted to Cohen's d effect size. 

### Female fitness data

These are for generation 39-41, from the file StandardFitnessG3941.csv. I ran a model following that used by the authors, as shown in their archived code.

```{r}
SF.data <- read_csv("input_data/lundhansen2020_github_data_FLXevoexp-master/StandardFitnessGeneration39-41 /StandardFitnessG3941.csv", show_col_types = F)
female_SF <- SF.data[SF.data$sex == "f", ] %>% 
  mutate(standard = as.numeric(scale(standard)))
male_SF <- SF.data[SF.data$sex == "m", ] %>% 
  mutate(standard = as.numeric(scale(standard)))

# control FM populations have non-sig higher female fitness than FLX populations (i.e. unexpected direction)
lund_females <- lm(standard ~ regime + rep_population, data = female_SF)

coef <- coefficients(summary(lund_females))

effects <- tibble(Estimate = coef[2,1],
                  Var = coef[2,2] ^ 2,
                  Lower_95_CI = coef[2,1] - coef[2,2]*1.96,
                  Upper_95_CI = coef[2,1] + coef[2,2]*1.96)


tibble(Study = "Lund-Hansen et al. (2020) Evolution 74, 2703–2713",
       Trait = "Female fitness",
       Generations = 41,
       Organism = "Drosophila melanogaster",
       Interpretation = "A positive effect would mean that females carrying female-evolved X chromosomes have higher fitness than the control FM genotype (there is no difference, though).",
       Effect_size_source = "Model of the raw data",
       Estimate = effects$Estimate,
       Var = effects$Var,
       Lower_95_CI = effects$Lower_95_CI,
       Upper_95_CI = effects$Upper_95_CI) %>% 
  write_csv("effect_sizes/lundhansen_1.csv")
```

### Male fitness

```{r}
# control FM populations have higher male fitness than FLX populations 
lund_males <- lm(standard ~ regime + rep_population, data = male_SF)

coef <- coefficients(summary(lund_males))

effects <- tibble(Estimate = coef[2,1],
                  Var = coef[2,2] ^ 2,
                  Lower_95_CI = coef[2,1] - coef[2,2]*1.96,
                  Upper_95_CI = coef[2,1] + coef[2,2]*1.96)


tibble(Study = "Lund-Hansen et al. (2020) Evolution 74, 2703–2713",
       Trait = "Male fitness",
       Generations = 41,
       Organism = "Drosophila melanogaster",
       Interpretation = "The positive effect indicates that males carrying female-evolved X chromosomes have lower fitness than the control FM genotype (there is no difference, though).",
       Effect_size_source = "Model of the raw data",
       Estimate = effects$Estimate,
       Var = effects$Var,
       Lower_95_CI = effects$Lower_95_CI,
       Upper_95_CI = effects$Upper_95_CI) %>% 
  write_csv("effect_sizes/lundhansen_2.csv")
```

### Female development time

```{r}
dev_time <- read_csv("input_data/lundhansen2020_github_data_FLXevoexp-master/DevelopmentTimeGeneration/DevelopmentTimeG43.csv", show_col_types = F)
dev_time_females <- dev_time %>% 
  filter(sex == "f") %>% 
  mutate(dev_time = as.numeric(scale(dev_time)))
dev_time_males <- dev_time %>% 
  filter(sex == "m") %>% 
  mutate(dev_time = as.numeric(scale(dev_time)))

# CFM females take non-sig longer to develop than FLX (this is the prediction, since shorter development is female-typical)
surv_obj <- Surv(dev_time_females$dev_time, event = rep(1, nrow(dev_time_females)))
cox <- coxph(surv_obj ~ regime, data = dev_time_females)

logHR <- coef(cox)["regimebCFM"]
Var_HR <- vcov(cox)["regimebCFM", "regimebCFM"]
CI_logHR <- logHR + c(-1, 1) * 1.96 * sqrt(Var_HR)

effects <- tibble(Estimate = logHR * sqrt(3) / pi,
                  Var = Var_HR * sqrt(3) / pi,
                  Lower_95_CI = CI_logHR[1] * sqrt(3) / pi,
                  Upper_95_CI = CI_logHR[2] * sqrt(3) / pi)

tibble(Study = "Lund-Hansen et al. (2020) Evolution 74, 2703–2713",
       Trait = "Female development time",
       Generations = 43,
       Organism = "Drosophila melanogaster",
       Interpretation = "A positive effect would mean that females carrying female-evolved X chromosomes take less time to develop higher fitness than the control FM genotype (there is no difference, though).",
       Effect_size_source = "Model of the raw data",
       Estimate = effects$Estimate,
       Var = effects$Var,
       Lower_95_CI = effects$Lower_95_CI,
       Upper_95_CI = effects$Upper_95_CI) %>% 
  write_csv("effect_sizes/lundhansen_3.csv")
```

### Male development time

```{r}
# CFM males take non-sig longer to develop than FLX (this is the prediction, since shorter development is a feminine trait)
surv_obj <- Surv(dev_time_males$dev_time, event = rep(1, nrow(dev_time_males)))
cox <- coxph(surv_obj ~ regime, data = dev_time_males)

logHR <- coef(cox)["regimebCFM"]
Var_HR <- vcov(cox)["regimebCFM", "regimebCFM"]
CI_logHR <- logHR + c(-1, 1) * 1.96 * sqrt(Var_HR)

effects <- tibble(Estimate = logHR * sqrt(3) / pi,
                  Var = Var_HR * sqrt(3) / pi,
                  Lower_95_CI = CI_logHR[1] * sqrt(3) / pi,
                  Upper_95_CI = CI_logHR[2] * sqrt(3) / pi)

tibble(Study = "Lund-Hansen et al. (2020) Evolution 74, 2703–2713",
       Trait = "Male development time",
       Generations = 43,
       Organism = "Drosophila melanogaster",
       Interpretation = "A positive effect would mean that males carrying female-evolved X chromosomes take less time to develop higher fitness than the control FM genotype (there is no difference, though).",
       Effect_size_source = "Model of the raw data",
       Estimate = effects$Estimate,
       Var = effects$Var,
       Lower_95_CI = effects$Lower_95_CI,
       Upper_95_CI = effects$Upper_95_CI) %>% 
  write_csv("effect_sizes/lundhansen_4.csv")
```

### Female locomotion

```{r warning=F}
loco <- read_csv("input_data/lundhansen2020_github_data_FLXevoexp-master/LocomotionAssayGeneration123/LocoAssayG123.csv", show_col_types = F)
female_loco <- loco %>% filter(sex == "f")
male_loco <- loco %>% filter(sex == "m")

# CFM > FLX for activity (on-sig), which is the prediction (so we need a positive effect size)
female_loco_model <- glm((cbind(active, inactive)) ~ regime + rep_population, 
                          data = female_loco, family = "binomial")

# calculate effect of regime on % active flies (as a log odds ratio) separately for each sex
LOR_effects <- contrast(emmeans(female_loco_model, ~ regime), method = "pairwise")[1,] %>% 
  as_tibble() %>% 
  rename(logOR = estimate) %>% 
  mutate(logOR = -1 * logOR) %>% 
  mutate(Estimate = logOR * sqrt(3) / pi,
         Var = (SE * sqrt(3) / pi)^2,
         Lower_95_CI = (logOR - SE * 1.96) * sqrt(3) / pi,
         Upper_95_CI = (logOR + SE * 1.96) * sqrt(3) / pi)

tibble(Study = "Lund-Hansen et al. (2020) Evolution 74, 2703–2713",
       Trait = "Female locomotion (2020)",
       Generations = 123,
       Organism = "Drosophila melanogaster",
       Interpretation = "A positive effect would mean that females carrying female-evolved X chromosomes are less active than females carrying the control FM genotype (there is no difference, though).",
       Effect_size_source = "Model of the raw data",
       Estimate = LOR_effects$Estimate,
       Var = LOR_effects$Var,
       Lower_95_CI = LOR_effects$Lower_95_CI,
       Upper_95_CI = LOR_effects$Upper_95_CI) %>% 
  write_csv("effect_sizes/lundhansen_5.csv")
```

### Male locomotion

```{r message=F}
# CFM < FLX for activity (and it's significant), which is the opposite of the prediction (so we need a negative effect size)
male_loco_model <- glm(cbind(active, inactive) ~ regime + rep_population, 
                          data = male_loco, family = "binomial")

# calculate effect of regime on % active flies (as a log odds ratio) separately for each sex
LOR_effects <- contrast(emmeans(male_loco_model, ~ regime), method = "pairwise")[1,] %>% 
  as_tibble() %>% 
  rename(logOR = estimate) %>% 
  mutate(logOR = -1 * logOR) %>% 
  mutate(Estimate = logOR * sqrt(3) / pi,
         Var = (SE * sqrt(3) / pi)^2,
         Lower_95_CI = (logOR - SE * 1.96) * sqrt(3) / pi,
         Upper_95_CI = (logOR + SE * 1.96) * sqrt(3) / pi)

tibble(Study = "Lund-Hansen et al. (2020) Evolution 74, 2703–2713",
       Trait = "Male locomotion (2020)",
       Generations = 123,
       Organism = "Drosophila melanogaster",
       Interpretation = "A positive effect would mean that males carrying female-evolved X chromosomes are less active than males carrying the control FM genotype (actually it's significant in the opposite direction).",
       Effect_size_source = "Model of the raw data",
       Estimate = LOR_effects$Estimate,
       Var = LOR_effects$Var,
       Lower_95_CI = LOR_effects$Lower_95_CI,
       Upper_95_CI = LOR_effects$Upper_95_CI) %>% 
  write_csv("effect_sizes/lundhansen_6.csv")
```

### Female thorax size

```{r}
thorax <- read_csv("input_data/lundhansen2020_github_data_FLXevoexp-master/ThoraxSizeGeneration72/ThoraxSizeG72.csv", show_col_types = F)
female_thorax <- thorax %>% 
  filter(sex == "f") %>% 
  mutate(size_mm = as.numeric(scale(size_mm)))
male_thorax <- thorax %>% 
  filter(sex == "m") %>% 
  mutate(size_mm = as.numeric(scale(size_mm)))

# FLX > CFM, which is predicted since large size is a feminine trait 
model_female_thorax <- lm(size_mm ~ regime + rep_population, data = female_thorax)

coef <- coefficients(summary(model_female_thorax))

effects <- tibble(Estimate = coef[2,1] * -1,
                  Var = coef[2,2] ^ 2,
                  Lower_95_CI = Estimate - coef[2,2]*1.96,
                  Upper_95_CI = Estimate + coef[2,2]*1.96)


tibble(Study = "Lund-Hansen et al. (2020) Evolution 74, 2703–2713",
       Trait = "Female thorax size",
       Generations = 72,
       Organism = "Drosophila melanogaster",
       Interpretation = "The positive effect indicates that females carrying female-evolved X chromosomes had larger thoraxes than females with the control FM genotype.",
       Effect_size_source = "Model of the raw data",
       Estimate = effects$Estimate,
       Var = effects$Var,
       Lower_95_CI = effects$Lower_95_CI,
       Upper_95_CI = effects$Upper_95_CI) %>% 
  write_csv("effect_sizes/lundhansen_7.csv")
```

### Male thorax size

```{r}
# FLX > CFM, which is predicted since large size is a feminine trait 
model_male_thorax <- lm(size_mm ~ regime + rep_population, data = male_thorax)

coef <- coefficients(summary(model_male_thorax))

effects <- tibble(Estimate = coef[2,1] * -1,
                  Var = coef[2,2] ^ 2,
                  Lower_95_CI = Estimate - coef[2,2]*1.96,
                  Upper_95_CI = Estimate + coef[2,2]*1.96)


tibble(Study = "Lund-Hansen et al. (2020) Evolution 74, 2703–2713",
       Trait = "Male thorax size",
       Generations = 72,
       Organism = "Drosophila melanogaster",
       Interpretation = "A positive effect would mean that males carrying female-evolved X chromosomes had larger thoraxes than males with the control FM genotype (there is no difference, though).",
       Effect_size_source = "Model of the raw data",
       Estimate = effects$Estimate,
       Var = effects$Var,
       Lower_95_CI = effects$Lower_95_CI,
       Upper_95_CI = effects$Upper_95_CI) %>% 
  write_csv("effect_sizes/lundhansen_8.csv")
```


## Manat et al. (2021)

Manat, Y., Lund-Hansen, K. K., Katsianis, G., & Abbott, J. K. (2021). Female-limited X-chromosome evolution effects on male pre- and post-copulatory success. _Biology Letters_, 17, 20200915.

### Male mating success

#### Manat et al.'s original data

The data were extracted from Figure 2a using PlotDigitizer. 

```{r}
df <- tibble(
  treatment = c("FLX", "FLX", "CFM", "CFM"),
  status = c("unmated", "mated", "unmated", "mated"),
  n = c(4,76,10,70)) %>% 
  pivot_wider(names_from = status, values_from = n)

kable(df) %>% 
  kable_styling(full_width = F)
```

#### Calculating effect size

```{r}
as.matrix(df[,2:3]) %>% 
  get_effect_size_2x2_matrix() %>% 
  mutate(Study = "Manat et al. (2021) Biol. Lett. 17, 20200915", 
         Trait = "Male mating success", 
         Generations = "107",
         Organism = "Drosophila melanogaster",
         Interpretation = "A positive effect size would mean that males carrying an X chromosome that underwent female-limited evolution would have lower mating success compared to males carrying control-evolved X chromosomes (however, there's a significant difference in the opposite direction).",
       Effect_size_source = "Chi-square test on total counts (pseudoreplicated)", 
         .before = 1) %>% 
  write_csv("effect_sizes/manat_1.csv")
```

### Mating speed

#### Manat et al.'s original data

The data were extracted from Figure 2b using PlotDigitizer. 

```{r}
df <- tibble(
  treatment = c("FLX", "CFM"),
  mean = c(10.14735444556975, 12.273534242844619),
  SE = c(11.07794959957106 - 10.14735444556975, 
         13.353237048825946 - 12.273534242844619),
  n = c(80,80))

kable(df) %>% 
  kable_styling(full_width = F)
```

#### Calculating effect size

```{r}
mes(m.1 = df$mean[1],
    m.2 = df$mean[2],
    sd.1 = df$SE[1] * sqrt(80),
    sd.2 = df$SE[2] * sqrt(80),
    n.1 = 880, n.2 = 80, verbose = F) %>% 
  parse_es(study = "Manat et al. (2021) Biol. Lett. 17, 20200915", 
           trait = "Male mating speed", 
           generations = "107",
           organism = "Drosophila melanogaster",
           interpretation = "A positive effect size would mean that males carrying an X chromosome that underwent female-limited evolution would take longer to induce females to mate compared to males carrying control-evolved X chromosomes (however, there is a significant difference in the opposite direction).") %>%  
  write_csv("effect_sizes/manat_2.csv")
```

### Male sperm defence

#### Manat et al.'s original data

The data were extracted from Figure 2c using PlotDigitizer. 

```{r}
df <- tibble(
  treatment = c("FLX", "CFM"),
  mean = c(0.21776979298517823, 0.1056360840036476),
  SE = c(0.26085500796210853 - 0.21776979298517823, 
         0.1243885508962476 - 0.1056360840036476),
  n = c(123,123))

kable(df) %>% 
  kable_styling(full_width = F)
```

#### Calculating effect size

```{r}
mes(m.1 = df$mean[2],
    m.2 = df$mean[1],
    sd.1 = df$SE[2] * sqrt(80),
    sd.2 = df$SE[1] * sqrt(80),
    n.1 = 880, n.2 = 80, verbose = F) %>% 
  parse_es(study = "Manat et al. (2021) Biol. Lett. 17, 20200915", 
           trait = "Sperm defence (2021)", 
           generations = "160",
           organism = "Drosophila melanogaster",
           interpretation = "A positive effect size would mean that males carrying an X chromosome that underwent female-limited evolution have worse sperm defence (i.e. sire fewer offspring after their mate mates with another male) compared to males carrying control-evolved X chromosomes (however, the finding is significant in the opposite direction).") %>% 
  write_csv("effect_sizes/manat_3.csv")
```

### Male sperm offence

#### Manat et al.'s original data

The data were extracted from Figure 2d using PlotDigitizer. 

```{r}
df <- tibble(
  treatment = c("FLX", "CFM"),
  mean = c(0.8606096836369432, 0.9267285480341434),
  SE = c(0.8815000797066794 - 0.8606096836369432, 
         0.9407380838514268 - 0.9267285480341434),
  n = c(123,123))

kable(df) %>% 
  kable_styling(full_width = F)
```

#### Calculating effect size

```{r}
mes(m.1 = df$mean[2],
    m.2 = df$mean[1],
    sd.1 = df$SE[2] * sqrt(80),
    sd.2 = df$SE[1] * sqrt(80),
    n.1 = 880, n.2 = 80, verbose = F) %>% 
  parse_es(study = "Manat et al. (2021) Biol. Lett. 17, 20200915", 
           trait = "Sperm defence (2021)", 
           generations = "160",
           organism = "Drosophila melanogaster",
           interpretation = "The positive effect indicates that males carrying an X chromosome that underwent female-limited evolution have worse sperm offence (i.e. sire fewer offspring with previously mated females) compared to males carrying control-evolved X chromosomes.") %>% 
  write_csv("effect_sizes/manat_4.csv")
```


## Lund-Hansen et al. (2022)

Lund-Hansen, K. K., Kutzer, M. A., Armitage, S. A., Gornard, S., Keilani, H., & Abbott, J. K. (2022). Female-limited X chromosome evolution reveals that lifespan is mainly modulated by interlocus rather than intralocus sexual conflict. _Behavioral Ecology and Sociobiology_, 76(9), 120.

### Female longevity 

These data come from Figure 1b, and consider female longevity/lifespan for females kept in mixed-sex groups. The raw data were provided by the authors in a Github repo ([link](https://github.com/KKLund-Hansen/FLXLongevityImmune)).

#### Calculating effect size

We calculated the effect size from a Cox model with fixed and random effects, re-using the archived analysis code to reproduce the exact model used by the authors. The model includes random effects accounting for repeated measures of vials and replicate populations in the experimental design, and includes treatment as a fixed effect. We then extracted the difference in survival rate between females carrying either a female-limited (FLX) or control X chromosome (specifically the CFM control, i.e. a control X that had co-evolved with X balancer chromosomes), expressed as a log hazard ratio, which we transformed to Cohen's $d$ by multiplying by $\sqrt{3}/\pi$. The effect size was set to negative, since females with the female-limited X showed non-significantly worse longevity (and thus possibly worse fitness, antagonistic pleiotropy notwithstanding) than controls.

```{r}
# Authors' original code start:
FLG.data <- read.csv(
  "input_data/LundHansen2022_FLXLongevityImmune/FemaleLongevity/FemaleLongevityGroup.csv", 
  header = T, sep = ",", stringsAsFactors = TRUE)
surv.FLG <- Surv(FLG.data$age, FLG.data$dead)
cox.FLG <- coxme(surv.FLG ~ regime + (1|regime/rep_population) + (1|vial), 
                 data = FLG.data)
# Authors' original code ends.

estimates <- contrast(emmeans(cox.FLG, "regime"), method = "pairwise") %>% 
  as.data.frame()

logHR <- estimates$estimate[estimates$contrast == "aFLX - bCFM"] * -1
Var_HR <- estimates$SE[estimates$contrast == "aFLX - bCFM"] ^ 2
CI_logHR <- logHR + c(-1, 1) * 1.96 * sqrt(Var_HR)

effects <- tibble(Estimate = logHR * sqrt(3) / pi,
                  Var = Var_HR * sqrt(3) / pi,
                  Lower_95_CI = CI_logHR[1] * sqrt(3) / pi,
                  Upper_95_CI = CI_logHR[2] * sqrt(3) / pi)

tibble(Study = "Lund-Hansen et al. (2022) Behav. Ecol. Sociobiol. 76, 120.",
       Trait = "Female longevity",
       Generations = 189,
       Organism = "Drosophila melanogaster",
       Interpretation = "A positive effect would mean that females carrying female-evolved X chromosomes survived longer in mixed-sex conditions than females with the control FM genotype (there is a non-significant effect in the opposite direction, though).",
       Effect_size_source = "Model of the raw data",
       Estimate = effects$Estimate,
       Var = effects$Var,
       Lower_95_CI = effects$Lower_95_CI,
       Upper_95_CI = effects$Upper_95_CI) %>% 
  write_csv("effect_sizes/lundhansen2022_1.csv")
```

### Male longevity 

These data come from Figure 1d, and consider male longevity/lifespan for males kept in mixed-sex groups. The raw data were provided by the authors in a Github repo ([link](https://github.com/KKLund-Hansen/FLXLongevityImmune)).

#### Calculating effect size

The statistical model and means of calculating effect size is the same as for the female longevity data. The effect size was positive, indicating males carrying a female-evolved X chromosome had (non-significantly) worse survival than males carrying a control-evolved X (as expected if the experimental removal of IaSC across the FLX allows male-harming alleles to increase).

```{r}
# Authors' original code start:
MLG.data <- read.csv(
  "input_data/LundHansen2022_FLXLongevityImmune/MaleLongevity/MaleLongevityGroup.csv", 
  header = T, sep = ",", stringsAsFactors = TRUE)
surv.MLG <- Surv(MLG.data$age, MLG.data$dead)
cox.MLG <- coxme(surv.MLG ~ regime + (1|regime/rep_population) + (1|vial), 
                 data = MLG.data)
# Authors' original code ends.

estimates <- contrast(emmeans(cox.MLG, "regime"), method = "pairwise") %>% 
  as.data.frame()

logHR <- estimates$estimate[estimates$contrast == "aFLX - bCFM"] 
Var_HR <- estimates$SE[estimates$contrast == "aFLX - bCFM"] ^ 2
CI_logHR <- logHR + c(-1, 1) * 1.96 * sqrt(Var_HR)

effects <- tibble(Estimate = logHR * sqrt(3) / pi,
                  Var = Var_HR * sqrt(3) / pi,
                  Lower_95_CI = CI_logHR[1] * sqrt(3) / pi,
                  Upper_95_CI = CI_logHR[2] * sqrt(3) / pi)

tibble(Study = "Lund-Hansen et al. (2022) Behav. Ecol. Sociobiol. 76, 120.",
       Trait = "Male longevity",
       Generations = 189,
       Organism = "Drosophila melanogaster",
       Interpretation = "A positive effect would mean that females carrying female-evolved X chromosomes survived longer in mixed-sex conditions than females with the control FM genotype (there is no difference, though).",
       Effect_size_source = "Model of the raw data",
       Estimate = effects$Estimate,
       Var = effects$Var,
       Lower_95_CI = effects$Lower_95_CI,
       Upper_95_CI = effects$Upper_95_CI) %>% 
  write_csv("effect_sizes/lundhansen2022_2.csv")
```

### Female bacterial resistance

These data come from the same Github repo ([link](https://github.com/KKLund-Hansen/FLXLongevityImmune)). The data give the mean bacterial load (in colony forming units, CFU) of the gram-negative _Lactococcus lactis_ in flies carrying three types of X chromosomes (female-evolved, control-evolved 'CFM', and wild type control). Higher bacterial load indicates lower resistance. 

#### Calculating effect size

We followed the authors' code , which involved transforming the bacterial load data via $log(x+1)$ and removing one outlier, then running a linear mixed model with random effects accounting for the experimental design. We standardised the response variable (after the log transformation) to have mean 0 and variance 1, such that the effect sizes were in units of Cohen's $d$.

```{r}
# Authors' code follows (except where noted!):
IA.data <- read.csv(
  "input_data/LundHansen2022_FLXLongevityImmune/InfectionAssay/infectionassay.csv", 
  header = T, sep = ",", stringsAsFactors = TRUE)

# log(x+1) transformation
IA.data$lnCFU <- log(IA.data$CFU_fly + 1)

# remove a major outlier
outlierIA <- IA.data$lnCFU[which((abs(IA.data$lnCFU - median(IA.data$lnCFU)) / mad(IA.data$lnCFU)) > 2)]
IA.data$lnCFU[IA.data$lnCFU %in% outlierIA[1]] <- NA

# This line added by Luke Holman: scaling response to mean 0, SD=1 to estimate Cohen's d
IA.data$lnCFU <- scale(IA.data$lnCFU)

#Linear mixed model testing if regime has a significant effect on bacterial load
model.BL <- lmer(lnCFU ~ regime + block + (1|regime:rep_population), data = IA.data)
# End of authors' code

estimates <- contrast(emmeans(model.BL, "regime"), method = "pairwise") %>% 
  as.data.frame()

d <- estimates$estimate[estimates$contrast == "aFLX - bCFM"] * -1
Var_d <- estimates$SE[estimates$contrast == "aFLX - bCFM"] ^ 2
CI_d <- d + c(-1, 1) * 1.96 * sqrt(Var_d)

tibble(Study = "Lund-Hansen et al. (2022) Behav. Ecol. Sociobiol. 76, 120.",
       Trait = "Female L. lactis resistance",
       Generations = 48,
       Organism = "Drosophila melanogaster",
       Interpretation = "A positive effect would mean that females carrying female-evolved X chromosomes were more resistant (i.e. had lower bacterial load of L. lactis) than females with the control FM genotype (there is no difference, though).",
       Effect_size_source = "Model of the raw data",
       Estimate = d,
       Var = Var_d,
       Lower_95_CI = CI_d[1],
       Upper_95_CI = CI_d[2]) %>% 
  write_csv("effect_sizes/lundhansen2022_3.csv")
```

<!-- ### Female bacterial tolerance -->

<!-- NOTE : LUKE REMOVED THIS DATA AS THE VARIANCE IS GIGANTIC AND IT MAKES THE PLOT LOOK WEIRD. THE META-ANALYSIS WOULD NOT BE AFFECTED SINCE THESE EFFECT SIZES HAVE HUGE UNCERTAINTY -->

<!-- These data come from the same Github repo ([link](https://github.com/KKLund-Hansen/FLXLongevityImmune)). The data give the number of surviving female flies 7 days post-treatment from experiments in which females were either naive, wounded, or infected with one of two bacteria (gram-negative _Lactococcus lactis_ or gram-positive _Pseudomonas entomophila_), and these females carried one of three types of X chromosomes (female-evolved, control-evolved 'CFM', and wild type control). Higher bacterial load indicates lower resistance. The wounding treatment is a procedural control for the mode of infection. -->

<!-- #### Calculating effect size -->

<!-- We re-used the authors' code to run the same Cox survival model with random effects. We then calculated interaction contrasts using `emmeans`, to 1) find the difference in survival probability between the wounding control and the bacteria treatments, and 2) compare this difference between the FLX and CFM treatments. Assuming removal of IaSC makes females healthier, we would predict a smaller worsening of survival in infected (vs wounded) flies in the FLX treatment compared to the CFM treatment. There is no statistically significant difference, however. -->

<!-- ```{r} -->
<!-- Sur7d.data <- read.csv("input_data/LundHansen2022_FLXLongevityImmune/SurvivalAfterInfection/survival7days.csv", h = T, sep = ",", stringsAsFactors = T) -->
<!-- surv.7d <- Surv(Sur7d.data$age, Sur7d.data$dead) -->
<!-- cox.7d <- coxme(surv.7d ~ regime * bacterium + block + (1|vial) + (1|regime/rep_population),  -->
<!--                 data = Sur7d.data) -->

<!-- # Calculate interaction contrasts from the author's model: -->

<!-- # Basically we want to know if the difference between: -->
<!-- ## A) the wounding control and the L. lactis treatment, and -->
<!-- ## B) the wounding control and the P. entomophila treatment -->
<!-- # differ between the FLX and CFM selection regimes. If yes, then the X chr treatment affects  -->
<!-- # survival tolerance to those two bacteria (in practice everything is non-significant) -->
<!-- bac_eff <- contrast( -->
<!--     emmeans(cox.7d, ~ bacterium | regime), -->
<!--     method = "trt.vs.ctrl", -->
<!--     ref = "bWounded"      -->
<!-- ) -->
<!-- x <- contrast(bac_eff, method = "pairwise", by = "contrast") %>%  -->
<!--   as.data.frame() -->

<!-- # Wounded vs L. lactis treatment -->
<!-- logHR_Llactis <- x$estimate[4] * -1 -->
<!-- varHR_Llactis <- x$SE[4] ^ 2 -->
<!-- CI_logHR_Llactis <- logHR_Llactis + c(-1, 1) * 1.96 * sqrt(varHR_Llactis) -->

<!-- # Wounded vs P. entomophila treatment -->
<!-- logHR_Pent<- x$estimate[7] * -1 -->
<!-- varHR_Pent <- x$SE[7] ^ 2 -->
<!-- CI_logHR_Pent <- logHR_Pent + c(-1, 1) * 1.96 * sqrt(varHR_Pent) -->

<!-- tibble(Study = "Lund-Hansen et al. (2022) Behav. Ecol. Sociobiol. 76, 120.", -->
<!--        Trait = "Female L. lactis tolerance", -->
<!--        Generations = 48, -->
<!--        Organism = "Drosophila melanogaster", -->
<!--        Interpretation = "A positive effect would mean that females carrying female-evolved X chromosomes were more tolerant of L. lactis infection (i.e. had a smaller reduction in survival 7 days after infection relative to the wounded control) than females with the control FM genotype (there is no difference, though).", -->
<!--        Effect_size_source = "Model of the raw data", -->
<!--        Estimate = logHR_Llactis, -->
<!--        Var = varHR_Llactis, -->
<!--        Lower_95_CI = CI_logHR_Llactis[1], -->
<!--        Upper_95_CI = CI_logHR_Llactis[2]) %>%  -->
<!--   write_csv("effect_sizes/lundhansen2022_4.csv") -->

<!-- tibble(Study = "Lund-Hansen et al. (2022) Behav. Ecol. Sociobiol. 76, 120.", -->
<!--        Trait = "Female P. entomophila tolerance", -->
<!--        Generations = 48, -->
<!--        Organism = "Drosophila melanogaster", -->
<!--        Interpretation = "A positive effect would mean that females carrying female-evolved X chromosomes were more tolerant of P. entomophila infection (i.e. had a smaller reduction in survival 7 days after infection relative to the wounded control) than females with the control FM genotype (there is no difference, though).", -->
<!--        Effect_size_source = "Model of the raw data", -->
<!--        Estimate = logHR_Pent, -->
<!--        Var = varHR_Pent, -->
<!--        Lower_95_CI = CI_logHR_Pent[1], -->
<!--        Upper_95_CI = CI_logHR_Pent[2]) %>%  -->
<!--   write_csv("effect_sizes/lundhansen2022_5.csv") -->
<!-- ``` -->

## Manat et al. (2026)

Manat, Y., Tarka, M., & Abbott, J. (in prep.) Change in the G-matrix as a result of female-limited X chromosome evolution in _Drosophila melanogaster_. _Philosophical Transactions of the Royal Society B: Biological Sciences_, in prep. (in the same Theme Issue as this meta-analysis).

```{r message=F, warning=F}
manat <- read_csv("input_data/Manat_PhilTransB_2026_Gmatrix/data.csv")
manat$round<-factor(manat$round)
manat$section<-factor(manat$section)
manat$round.sec<-factor(manat$round.sec)

females <- subset(manat, manat$sex=="female")
males <- subset(manat, manat$sex=="male")

female_DR <- aggregate(females$DR, 
                       by=list(mother=females$mother, 
                               replicat=females$replicat, 
                               treatment=females$treatment, 
                               round=females$round), FUN=mean, na.action=na.rm) %>% 
  filter(!is.na(x)) %>% 
  mutate(x = scale(x))

male_DR <- aggregate(males$DR, 
                     by=list(mother=males$mother, 
                             replicat=males$replicat, 
                             treatment=males$treatment, 
                             round=males$round), 
                     FUN=mean, na.action=na.rm) %>% 
  filter(!is.na(x)) %>% 
  mutate(x = scale(x))

female_WS <- aggregate(females$WS, 
                       by=list(mother=females$mother,
                               replicat=females$replicat,
                               treatment=females$treatment,
                               round=females$round), 
                       FUN=mean, na.action=na.rm) %>% 
  filter(!is.na(x)) %>% 
  mutate(x = scale(x))

male_WS <- aggregate(males$WS, 
                     by=list(mother=males$mother,
                             replicat=males$replicat,
                             treatment=males$treatment,
                             round=males$round), 
                     FUN=mean, na.action=na.rm) %>% 
  filter(!is.na(x)) %>% 
  mutate(x = scale(x))

female_LA <- aggregate(females$LA_active, 
                 by=list(mother=females$mother,
                         replicat=females$replicat,
                         treatment=females$treatment,
                         round=females$round), FUN=sum) %>% 
  filter(!is.na(x)) 

male_LA <- aggregate(males$LA_active, 
                     by=list(mother=males$mother,
                             replicat=males$replicat,
                             treatment=males$treatment,
                             round=males$round), FUN=sum) %>% 
  filter(!is.na(x))


modelF.dr <- lmer(x ~ treatment + round + (1 | treatment:replicat), data = female_DR)
modelM.dr <- lmer(x ~ treatment + round + (1 | treatment:replicat), data = male_DR)
modelF.ws<- lmer(x~treatment+round+(1|treatment:replicat), data=female_WS)
modelM.ws<- lmer(x~treatment+round+(1|treatment:replicat), data=male_WS)
modelF.la<- glmer(x~treatment+round+(1|treatment:replicat), data=female_LA, family=poisson)
modelM.la<- glmer(x~treatment+round+(1|treatment:replicat), data=male_LA, family=poisson)

effects <- tibble(trait = c("Female dessication resistance", "Male dessication resistance", 
                            "Female wing size", "Male wing size", 
                            "Female locomotion", "Male locomotion"),
       rbind(
         as.data.frame(contrast(emmeans(modelF.dr, ~ treatment), method = "pairwise"))[2,],
         as.data.frame(contrast(emmeans(modelM.dr, ~ treatment), method = "pairwise"))[2,],
         as.data.frame(contrast(emmeans(modelF.ws, ~ treatment), method = "pairwise"))[2,],
         as.data.frame(contrast(emmeans(modelM.ws, ~ treatment), method = "pairwise"))[2,],
         as.data.frame(contrast(emmeans(modelF.la, ~ treatment), method = "pairwise"))[2,],
         as.data.frame(contrast(emmeans(modelM.la, ~ treatment), method = "pairwise"))[2,]
         )) 

# flip sign of dessication result for females, since authors say it's opposite to prediction in their Abstract
# I also flip the sign for the wings results, since we want these effects to be positive given FLX are larger (feminised), just non-sig so
effects$estimate[1:4] <- effects$estimate[1:4] * -1 
effects <- effects %>% 
  mutate(Var = SE ^ 2,
         Lower_95_CI = estimate - 1.96 * SE,
         Upper_95_CI = estimate + 1.96 * SE)
  
  
effects[5:6, ] <- 
  effects[5:6, ] %>% 
   mutate(Var = (SE * sqrt(3) / pi)^2,
         Lower_95_CI = (estimate - SE * 1.96) * sqrt(3) / pi,
         Upper_95_CI = (estimate + SE * 1.96) * sqrt(3) / pi,
         estimate = estimate * sqrt(3) / pi)

tibble(Study = "Manat et al. (2026) Phil. Trans. B, in prep.",
       Trait = effects$trait,
       Generations = 147,
       Organism = "Drosophila melanogaster",
       Interpretation = c(
         "A positive effect size would mean that females carrying a female-limited X show increased (i.e. female-like) dessication resistance relative to controls. Actually, they showed significantly lower resistance than controls.",
         "A positive effect size would mean that males carrying a female-limited X show increased (i.e. female-like) dessication resistance relative to controls. Actually, they showed non-significantly lower resistance than controls.",
         "A positive effect size would mean that females carrying a female-limited X show increased (i.e. female-like) wing size relative to controls. Actually, they had non-significantly larger wings than controls.",
         "A positive effect size would mean that males carrying a female-limited X show increased (i.e. female-like) wing size relative to controls. Actually, they had non-significantly larger wings than controls.",
         "A positive effect size would mean that females carrying a female-limited X show reduced (i.e. female-like) locomotion relative to controls. Actually, they showed non-significantly higher locomotion than controls.",
         "A positive effect size would mean that males carrying a female-limited X show reduced (i.e. female-like) locomotion relative to controls. The treatment was in the predicted direction but was non-significant."),
       Effect_size_source = "Model of the raw data",
       Estimate = effects$estimate,
       Var = effects$Var,
       Lower_95_CI = effects$Lower_95_CI,
       Upper_95_CI = effects$Upper_95_CI) %>% 
  write_csv("effect_sizes/manat_phil_trans.csv")
```

## Lund-Hansen et al. (2026)


Lund-Hansen, K. K., Li, Q. Y., Goyal, V., & Abbott, J. K. (in prep.) Signatures of sexually antagonistic intragenomic coevolution in _Drosophila melanogaster_. _Philosophical Transactions of the Royal Society B: Biological Sciences_, in prep. (in the same Theme Issue as this meta-analysis).

### Locomotory activity

```{r}
Locoswap.data <- read.csv(file = "input_data/LundHansen_PhilTransB_2026/locomotionswap.csv", h = T, sep = ",", stringsAsFactors = T)

model_loco <- glmer(cbind(active, inactive) ~ type * sex + day + (1|type:rep_pop) + (1|time), data = Locoswap.data, family = binomial)

effects <- contrast(emmeans(model_loco, ~ type | sex), method = "pairwise") %>% 
  as.data.frame() %>% as_tibble() %>% 
  filter(contrast == "aFLXsex - cCFMsex") %>% 
  mutate(LOR = -1 * estimate) %>% # flip effect sign to positive 
  mutate(Estimate = LOR * sqrt(3) / pi, # convert LOR to Cohen's d
         Var = (SE * sqrt(3) / pi)^2,
         Lower_95_CI = (LOR - SE * 1.96) * sqrt(3) / pi,
         Upper_95_CI = (LOR + SE * 1.96) * sqrt(3) / pi)

tibble(Study = "Lund-Hansen et al. (2026) Phil. Trans. B, in prep.",
       Trait = c("Female locomotion", "Male locomotion"),
       Generations = 169,
       Organism = "Drosophila melanogaster",
       Interpretation = c(
         "The positive effect size indicates that females carrying a female-limited X and Cwt autosomes show reduced (i.e. female-like) locomotion relative to females carrying a control (CFM) X and Cwt autosomes.",
         "A positive effect size would indicate that males carrying a female-limited X and Cwt autosomes show reduced (i.e. female-like) locomotion relative to females carrying a control (CFM) X and Cwt autosomes (it't non-significant, though)."),
       Effect_size_source = "Model of the raw data",
       Estimate = effects$Estimate,
       Var = effects$Var,
       Lower_95_CI = effects$Lower_95_CI,
       Upper_95_CI = effects$Upper_95_CI) %>% 
  write_csv("effect_sizes/lundhansen_phil_trans1.csv")
```

### Sperm defence

```{r}
defswap.data <- read.csv(file = "input_data/LundHansen_PhilTransB_2026/defenceswap.csv", h = T, sep = ",", stringsAsFactors = T)

model_def <- glmer(cbind(red, brown) ~ type + block + (1|type:rep_pop), data = defswap.data, family = binomial)

effects <- contrast(emmeans(model_def, ~ type ), method = "pairwise") %>% 
  as.data.frame() %>% as_tibble() %>% 
  filter(contrast == "aFLXsex - cCFMsex") %>% 
  mutate(LOR = estimate * -1) %>% # we want negative sign, since FLX males were non-sig. better competitors (opposite to prediction)
  mutate(Estimate = LOR * sqrt(3) / pi, # convert LOR to Cohen's d
         Var = (SE * sqrt(3) / pi)^2,
         Lower_95_CI = (LOR - SE * 1.96) * sqrt(3) / pi,
         Upper_95_CI = (LOR + SE * 1.96) * sqrt(3) / pi)


tibble(Study = "Lund-Hansen et al. (2026) Phil. Trans. B, in prep.",
       Trait = "Sperm defence",
       Generations = 169,
       Organism = "Drosophila melanogaster",
       Interpretation = "A positive effect size would indicate that males carrying a female-limited X and Cwt autosomes show worse sperm defence than males carrying a control (CFM) X and Cwt autosomes (there is no significant difference, however).",
       Effect_size_source = "Model of the raw data",
       Estimate = effects$Estimate,
       Var = effects$Var,
       Lower_95_CI = effects$Lower_95_CI,
       Upper_95_CI = effects$Upper_95_CI) %>% 
  write_csv("effect_sizes/lundhansen_phil_trans2.csv")
```

### Sperm offence

```{r}
offswap.data <- read.csv(file = "input_data/LundHansen_PhilTransB_2026/offenceswap.csv", h = T, sep = ",", stringsAsFactors = T)

model_off <- glmer(cbind(red, brown) ~ type + block + (1|type:rep_pop), data = offswap.data, family = binomial)

effects <- contrast(emmeans(model_off, ~ type ), method = "pairwise") %>% 
  as.data.frame() %>% as_tibble() %>% 
  filter(contrast == "aFLXsex - cCFMsex") %>% 
  mutate(LOR = estimate) %>% 
  mutate(Estimate = LOR * sqrt(3) / pi, # convert LOR to Cohen's d
         Var = (SE * sqrt(3) / pi)^2,
         Lower_95_CI = (LOR - SE * 1.96) * sqrt(3) / pi,
         Upper_95_CI = (LOR + SE * 1.96) * sqrt(3) / pi)


tibble(Study = "Lund-Hansen et al. (2026) Phil. Trans. B, in prep.",
       Trait = "Sperm offence",
       Generations = 169,
       Organism = "Drosophila melanogaster",
       Interpretation = "A positive effect size would indicate that males carrying a female-limited X and Cwt autosomes show worse sperm offence than males carrying a control (CFM) X and Cwt autosomes (there is no significant difference, however).",
       Effect_size_source = "Model of the raw data",
       Estimate = effects$Estimate,
       Var = effects$Var,
       Lower_95_CI = effects$Lower_95_CI,
       Upper_95_CI = effects$Upper_95_CI) %>% 
  write_csv("effect_sizes/lundhansen_phil_trans3.csv")
```



## Grieshop et al. (2025)

Grieshop, K., Liu, M. J., Frost, R. S., Lindsay, M. P., Bayoumi, M., Brengdahl, M. I., Molnar, R. I., & Agrawal, A. F. (2025). Expression divergence in response to sex-biased selection. _Molecular Biology and Evolution_, 42(7) msaf099.

Here, I obtained the raw data from the authors, and re-ran a binomial GLMM that followed the original analysis as closely as I could figure out. The model measures the probability of mating (vs not mating) for males and females and flies carrying 'Red' chromosomes (copies of Chromosome 2 which spent most of their recent evolutionary history in males) versus 'Non-Red' chromosomes (which spent a majority of generations in females). From the binomial GLM I get the treatment effect size as a log odds ratio, and then convert that to Cohen's d.

```{r}
###### This block of code copied from the authors' archived R script ########################################################################

#Load in fitness data
DsRed_Data <- read.csv("input_data/grieshop2025_DsRed_Data_Formatted.csv")

# Data Prep
#######
# Set up a vector full of zeroes of the appropriate length, then...
Prop_Mated <- rep.int(0, dim(DsRed_Data)[1])

# iterate over each observation in the DsRed_Data object
for (i in 1:dim(DsRed_Data)[1]){
  # if the observation is Male for that population replicate cage and generation
  if (DsRed_Data$Sex[i] == "Male"){
    # the proportion of mated males 
    Prop_Mated[i] <- DsRed_Data$Male_Prop_Mated[i]
  }
  else{
    # the proportion of mated females for that population replicate cage and generation
    Prop_Mated[i] <- DsRed_Data$Female_Prop_Mated[i]
  }
  #... fill in said vector at any given point with the appropriate value 
  # for the proportion of sampled individuals of the given sex found to be mated...
}

#...and then add that vector as another column.
DsRed_Data <- cbind(DsRed_Data, Prop_Mated) 

markerXpop.bothsex.model <- glmer(
  Mating_Status == "Mated" ~ Marker*Sex + (Marker|Population),
  binomial(link='probit'), offset = Prop_Mated, data = DsRed_Data)

###### End of copied code ####################################

# calculate effect of marker on mating rate (as a log odds ratio) separately for each sex
LOR_effects <- contrast(emmeans(markerXpop.bothsex.model, ~ Marker | Sex), method = "pairwise") %>% 
  as_tibble() %>% 
  mutate(estimate = c(1,-1) * estimate) %>% # change sign of the effect size for males
  rename(logOR = estimate) %>% 
  mutate(Estimate = logOR * sqrt(3) / pi, # convert LOR to Cohen's d
         Var = (SE * sqrt(3) / pi)^2,
         Lower_95_CI = (logOR - SE * 1.96) * sqrt(3) / pi,
         Upper_95_CI = (logOR + SE * 1.96) * sqrt(3) / pi)

tibble(Study = "Grieshop et al. (2025) Molec. Biol. Evol. 42 msaf099.",
       Trait = c("Female mating rate", "Male mating rate"),
       Generations = 103,
       Organism = "Drosophila melanogaster",
       Interpretation = c(
         "A positive effect would mean that females carrying primarily male-evolved copies of chromosome 2 mate less often, compared to females with primarily female-evolved copies of chromosome 2 (there is no difference, though).", 
         "This positive effect means that males carrying primarily male-evolved copies of chromosome 2 mate more often, compared to males with primarily female-evolved copies of chromosome 2."),
       Effect_size_source = "Model of the raw data",
       Estimate = LOR_effects$Estimate,
       Var = LOR_effects$Var,
       Lower_95_CI = LOR_effects$Lower_95_CI,
       Upper_95_CI = LOR_effects$Upper_95_CI) %>% 
  write_csv("effect_sizes/grieshop2025.csv")
```


## Melo-Gavin et al. (2025)

Melo-Gavin, C., Lindsay, M. P., & Agrawal, A. F. (2025). Response of mate harm to sex-separated gene pools. bioRxiv, 2025-09.

### Melo-Gavin et al.'s raw data

The data were obtained from the left two bars in Figure 1, using PlotDigitizer. The figure shows means and 95% CIs, so I calculated the standard deviation from the CIs as `(CI_95 / 1.96) * sqrt(N)`, i.e. assuming the CIs were calculated as 1.96 times the standard error of the mean.

```{r}
df <- tibble(male_type = c("NonRed", "Red"),
       line_type = c("Treatment", "Treatment"),
       mean = c(57.18512618727785, 53.99935806754556),
       CI_95 = c(59.0035781790516 - 57.18512618727785,
                 55.64306178005492 - 53.99935806754556),
       N = c(180, 180),
       SE = (CI_95 / 1.96),
       SD = SE * sqrt(N))

mes(m.1 = df$mean[df$male_type=="NonRed"],
    m.2 = df$mean[df$male_type=="Red"],
    sd.1 = df$SD[df$male_type=="NonRed"],
    sd.2 = df$SD[df$male_type=="Red"],
    n.1 = 180, n.2 = 180, verbose=F) %>% 
  parse_es(study = "Melo-Gavin et al. (2025) bioRxiv, https://doi.org/10.1101/2025.09.29.678620.",
           generations = 11,
           organism = "Drosophila melanogaster",
           trait = "Male harm to females",
           interpretation = "The positive effect indicates that males carrying male-evolved chromosomes reduced the fecundity of their female mates more than did males carrying chromosomes that had undergone female-biased evolution."
          ) %>% 
  write_csv("effect_sizes/melo_gavin2025_maleharm.csv")
```

## Thyagarajan et al. (2025)

Thyagarajan, H., Sayyed, I., Baroody, M. G., Kowal, J. A., Day, T., & Chippindale, A. K. (2025). Mixed evidence for intralocus sexual conflict from male-limited selection in _Drosophila melanogaster_. _Journal of Heredity_, esaf072.

### Female and male fitness

These variables are referred to as 'competitive reproductive fitness' in the paper. I obtained the raw data from Dryad and ran one model for each sex. The models were binomial GLMs with treatment (male-evolved or control), replicate, and generation as fixed effects (no interactions are fitted, i.e. I estimate the average treatment effect across the 3 starting genomes, labelled 'replicates'). From the binomial GLM I get the treatment effect size as a log odds ratio, and then convert that to Cohen's d. For both sexes, the effect size compares the control and single-dose male limited treatments.

```{r}
female_CRF <- read_csv("input_data/Thyagarajan_2025_dryad_data/F_CRF_wt.cr.csv",
                       show_col_types = F) %>% 
  mutate(Sex = "F", Replicate = factor(Replicate))
male_CRF <- read_csv("input_data/Thyagarajan_2025_dryad_data/M_CRF_wt.cr.csv",
                     show_col_types = F) %>% 
  mutate(Sex = "M", Replicate = factor(Replicate)) %>% 
  filter(Treatment %in% c("MLSD", "MC"))


# calculate effect of marker on mating rate (as a log odds ratio) separately for each sex
## GLM for females:
model <- glm(cbind(Red, Brown) ~ Treatment + Replicate + Gen, data = female_CRF, family = "quasibinomial")
female_effects <- contrast(emmeans(model, ~ Treatment), method = "pairwise") %>% 
  as_tibble() %>% 
  rename(logOR = estimate) %>% 
  mutate(Estimate = logOR * sqrt(3) / pi,
         Var = (SE * sqrt(3) / pi)^2,
         Lower_95_CI = (logOR - SE * 1.96) * sqrt(3) / pi,
         Upper_95_CI = (logOR + SE * 1.96) * sqrt(3) / pi)

## GLM for males:
model2 <- glm(cbind(Red, Brown) ~ Treatment + Replicate + factor(Gen), data = male_CRF, family = "quasibinomial")
male_effects <- contrast(emmeans(model2, ~ Treatment), method = "pairwise") %>% 
  as_tibble() %>% 
  mutate(estimate = estimate * -1) %>% # change sign of effect size (it goes in the predicted direction)
  rename(logOR = estimate) %>% 
  mutate(Estimate = logOR * sqrt(3) / pi,
         Var = (SE * sqrt(3) / pi)^2,
         Lower_95_CI = (logOR - SE * 1.96) * sqrt(3) / pi,
         Upper_95_CI = (logOR + SE * 1.96) * sqrt(3) / pi)

LOR_effects <- bind_rows(female_effects, male_effects)

tibble(Study = "Thyagarajan et al. (2025) J. Hered. esaf072",
       Trait = c("Female fitness", "Male fitness"),
       Generations = c(50, 64),
       Organism = "Drosophila melanogaster",
       Interpretation = c("This positive effect means that females carrying male-evolved haplotypes have lower fitness than females carrying control genomes.", 
                          "This positive effect means that males carrying male-evolved haplotypes have higher fitness than males carrying control genomes."),
       Effect_size_source = "Model of the raw data",
       Estimate = LOR_effects$Estimate,
       Var = LOR_effects$Var,
       Lower_95_CI = LOR_effects$Lower_95_CI,
       Upper_95_CI = LOR_effects$Upper_95_CI) %>% 
  write_csv("effect_sizes/thyagarajan2025_1.csv")

```

### Male mating success

```{r message=F}
mating_competition <- read_csv("input_data/Thyagarajan_2025_dryad_data/M_MC_wt.cr.csv",
                               show_col_types = F) %>% 
  filter(!is.na(Mated.male)) %>% 
  group_by(Selection, Replicate) %>% 
  summarise(n_wt = sum(Mated.male == "wt"),
            n_bw = sum(Mated.male == "bw"))

model <- glm(cbind(n_wt, n_bw) ~ Selection + Replicate, data = mating_competition, family = "binomial")

LOR_effects <- contrast(emmeans(model, ~ Selection), method = "pairwise") %>% 
  as_tibble() %>% 
  mutate(estimate = estimate * -1) %>% # change sign of effect size (it goes in the predicted direction)
  rename(logOR = estimate) %>% 
  mutate(Estimate = logOR * sqrt(3) / pi,
         Var = (SE * sqrt(3) / pi)^2,
         Lower_95_CI = (logOR - SE * 1.96) * sqrt(3) / pi,
         Upper_95_CI = (logOR + SE * 1.96) * sqrt(3) / pi)

tibble(
  Study = "Thyagarajan et al. (2025) J. Hered. esaf072",
  Trait = "Male mating success",
  Generations = 64,
  Organism = "Drosophila melanogaster",
  Interpretation = "A positive effect would mean that males carrying male-evolved haplotypes are more likely to mate with females (over standard male competitors) compared to males carrying control genomes (there is no signficant difference).",
  Effect_size_source = "Model of the raw data",
  Estimate = LOR_effects$Estimate,
  Var = LOR_effects$Var,
  Lower_95_CI = LOR_effects$Lower_95_CI,
  Upper_95_CI = LOR_effects$Upper_95_CI) %>% 
  write_csv("effect_sizes/thyagarajan2025_2.csv")
```


### Male sperm offence

```{r}
sperm_comp_data <- read_csv("input_data/Thyagarajan_2025_dryad_data/M_SC_wt.cr.csv",
                            show_col_types = F) %>% 
  filter(Treatment %in% c("MC", "MLSD")) %>% 
  mutate(Replicate = factor(Replicate))

model <- glm(cbind(WT, U) ~ Treatment + Replicate, family = "binomial", data = sperm_comp_data)
LOR_effects <- contrast(emmeans(model, ~ Treatment), method = "pairwise") %>% 
  as_tibble() %>% 
  mutate(estimate = estimate * -1) %>% # change sign of effect size (it goes in the predicted direction)
  rename(logOR = estimate) %>% 
  mutate(Estimate = logOR * sqrt(3) / pi,
         Var = (SE * sqrt(3) / pi)^2,
         Lower_95_CI = (logOR - SE * 1.96) * sqrt(3) / pi,
         Upper_95_CI = (logOR + SE * 1.96) * sqrt(3) / pi)

tibble(
  Study = "Thyagarajan et al. (2025) J. Hered. esaf072",
  Trait = "Sperm offence",
  Generations = 64,
  Organism = "Drosophila melanogaster",
  Interpretation = "A positive effect would indicate that males carrying male-evolved haplotypes sire more offspring when mating with previously-mated females, compared to males carrying control genomes (there is no significant effect).",
  Effect_size_source = "Model of the raw data",
  Estimate = LOR_effects$Estimate,
  Var = LOR_effects$Var,
  Lower_95_CI = LOR_effects$Lower_95_CI,
  Upper_95_CI = LOR_effects$Upper_95_CI) %>% 
  write_csv("effect_sizes/thyagarajan2025_3.csv")
```

### Female productivity

```{r}
female_prod_data <- read_csv("input_data/Thyagarajan_2025_dryad_data/F_HClines.csv",
                            show_col_types = F) %>% 
  mutate(Pupal.count = as.numeric(scale(Pupal.count)),
         Dams.Count = as.numeric(scale(Dams.Count))) %>% 
  mutate(Line = factor(Line),
         Vial = paste(Line, Vial, sep = "_"))

model <- lmer(Pupal.count ~ Dams.Count + Selection + (1| Line), data = female_prod_data) 

effects <- contrast(emmeans(model, ~ Selection), method = "pairwise") %>% 
  as_tibble() %>% 
  mutate(Estimate = estimate,
         Var = SE ^ 2,
         Lower_95_CI = (Estimate - SE * 1.96),
         Upper_95_CI = (Estimate + SE * 1.96))

tibble(
  Study = "Thyagarajan et al. (2025) J. Hered. esaf072",
  Trait = "Female productivity",
  Generations = 64,
  Organism = "Drosophila melanogaster",
  Interpretation = "A positive effect would indicate that females carrying male-evolved haplotypes produce fewer offspring compared to females carrying control genomes (there is a non-significant trend in the opposite direction).",
  Effect_size_source = "Model of the raw data",
  Estimate = effects$Estimate,
  Var = effects$Var,
  Lower_95_CI = effects$Lower_95_CI,
  Upper_95_CI = effects$Upper_95_CI) %>% 
  write_csv("effect_sizes/thyagarajan2025_4.csv")
```




## Nordén et al. (2023) 

Nordén, A. K., Ramm, S. A., & Abbott, J. K. (2023). Rapid evolution of sex role specialization in a hermaphrodite under sex-limited selection. _Evolution_, 77(4), 1066-1076.

The raw data were taken from the Dryad repository, and analysed using the same model as in the original study (copying the R code). The fitness data were mean-centered and scaled prior to analysis, such that model-estimated differences in treatment means are equal to Cohen's d. 

Note that this study has both a female-limited and a male-limited evolution treatment, unlike all the others in this section, and it measured both male and female fitness. To address this, I include 4 contrasts in the meta-analysis: female-limited evolution vs the control for both sexes' fitnesses, and male-limited evolution vs the control for both sexes' fitnesses. I predict elevated fitness in the selected sex and reduced fitness in the non-selected sex, and adjust the effect size signs to reflect this.

```{r}
fitness.dataCS <- read.csv("input_data/norden_dryad_data/Mlig_fitness_data_C_standard.csv", 
                           sep=";", dec=",", header=TRUE, stringsAsFactors = T) %>% 
  as_tibble() %>% 
  mutate(Stand_Fitness = as.numeric(scale(Stand_Fitness)))

modfitCS <- lmer(Stand_Fitness~Treatment*sex_role + (1|ID) + (1|Treatment:sex_role:Population), data = fitness.dataCS)

cohens_d <- emmeans(modfitCS, ~ Treatment | sex_role) %>% 
  pairs() %>% as_tibble() %>% 
  filter(str_detect(contrast, "Control")) %>% 
  select(trait = contrast, Estimate = estimate, SE)

cohens_d$trait <- c(
  "Female fitness (female-limited lines)",
  "Female fitness (male-limited lines)",
  "Male fitness (female-limited lines)",
  "Male fitness (male-limited lines)")

cohens_d$Estimate <- cohens_d$Estimate * c(-1, 1, 1, -1) 

cohens_d$Lower_95_CI <- cohens_d$Estimate - cohens_d$SE * 1.96
cohens_d$Upper_95_CI <- cohens_d$Estimate + cohens_d$SE * 1.96

tibble(Study = "Nordén et al. (2020) Evolution 77, 1066-1076",
       Trait = cohens_d$trait,
       Generations = 14,
       Organism = "Macrostomum lignano",
       Interpretation = c(
         "The positive effect size means that worms carrying the female-limited genomic region had higher fitness when breeding in the female sex role compared to worms carrying the control allele.",
         "A positive effect would mean that worms carrying the male-limited genomic region had lower fitness when breeding in the female sex role compared to worms carrying the control allele (actually, they had non-significantly higher fitness).",
         "A positive effect would mean that worms carrying the female-limited genomic region had lower fitness when breeding in the male sex role compared to worms carrying the control allele (they did, but the difference is not significant).",
         "A positive effect would mean that worms carrying the male-limited genomic region had higher fitness when breeding in the male sex role compared to worms carrying the control allele (they did, but the difference is not significant)."
       ),
       Effect_size_source = "Model of the raw data",
       Estimate = cohens_d$Estimate,
       Var = cohens_d$SE ^ 2,
       Lower_95_CI = cohens_d$Lower_95_CI,
       Upper_95_CI = cohens_d$Upper_95_CI) %>% 
  write_csv("effect_sizes/norden2020.csv")
```

## Cīrulis et al. (2026)

Cīrulis, A., Majvall, M., & Abbott, J. K. (2026). Sex-limited experimental evolution, mating, and sexual conflict in a hermaphrodite. _Zoological Journal of the Linnean Society_, 206(3), zlag036.

### Mating frequency

#### Calculation of effect size

The raw data appear as a supplement to the paper on the journal webpages. The authors write (in legend for Table S2) they analysed these data with a "linear mixed effects model with selection regime as a fixed effect, replicate population as a random effect nested within selection regime, and generation as a random effect, to test if mating frequency differs between selection regimes." I have reproduced this model in the code below, except that I first transformed the response to mean 0 and unit variance, in order to find Cohen's d using planned contrasts (between the control and F treatment, and control and M treatment).

```{r}
cirulis_dat1 <- read_xlsx("input_data/cirulis_2026_archived_data/Data-1.mating_experiment cutoff - 3(G41,43,44).xlsx") %>% 
  mutate(scaled_matings = scale(Matings))

mod <- lmer(scaled_matings ~ Treatment + (1|Treatment:Population) + (1|Generation), 
            data = cirulis_dat1)
effects <- contrast(emmeans(mod, "Treatment"), method = "pairwise") %>% 
  as.data.frame()

F_effects <- tibble(
  Estimate = effects$estimate[effects$contrast == "C - F"],
  SE = effects$SE[effects$contrast == "C - F"],
  Lower_95_CI = Estimate - SE*1.96,
  Upper_95_CI = Estimate+ SE*1.96)

M_effects <- tibble(
  Estimate = effects$estimate[effects$contrast == "C - M"] * -1,
  SE = effects$SE[effects$contrast == "C - M"],
  Lower_95_CI = Estimate - SE*1.96,
  Upper_95_CI = Estimate + SE*1.96)

tibble(Study = "Cīrulis et al. (2026). Sex-limited experimental evolution, mating, and sexual conflict in a hermaphrodite. Zoo J Linnean Soc, 206, zlag036.",
       Trait = "Mating rate (female-limited lines)",
       Generations = 44,
       Organism = "Macrostomum lignano",
       Interpretation = "A positive effect size would indicate that worms carrying the female-limited genomic region show reduced mating behaviour (it's non-significant, though).",
       Effect_size_source = "Model of the raw data",
       Estimate = F_effects$Estimate,
       Var = F_effects$SE ^ 2,
       Lower_95_CI = F_effects$Lower_95_CI,
       Upper_95_CI = F_effects$Upper_95_CI) %>% 
  write_csv("effect_sizes/cirulis2026_1.csv")

tibble(Study = "Cīrulis et al. (2026). Sex-limited experimental evolution, mating, and sexual conflict in a hermaphrodite. Zoo J Linnean Soc, 206, zlag036.",
       Trait = "Mating rate (male-limited lines)",
       Generations = 44,
       Organism = "Macrostomum lignano",
       Interpretation = "A positive effect size would indicate that worms carrying the male-limited genomic region show increased mating behaviour (it's non-significant, though).",
       Effect_size_source = "Model of the raw data",
       Estimate = M_effects$Estimate,
       Var = M_effects$SE ^ 2,
       Lower_95_CI = M_effects$Lower_95_CI,
       Upper_95_CI = M_effects$Upper_95_CI) %>% 
  write_csv("effect_sizes/cirulis2026_2.csv")
```

### Effect on mate's behaviour 

The authors measured whether or not wild type worms performed a post-mating behaviour called 'sucking' after having mated to selection line worms (either male-evolved, female-evolved, or control lines). The sucking behaviour is performed by a worm on itself after mating, and is interpreted as an attempt to remove the sperm from the mating partner (i.e. an inter-locus sexual conflict over mating role in this hermaphroditic species). 

#### Source of the data

The authors kindly provided the relevant data in a separate file, which is loaded here.

```{r message=F}
cirulis_dat2 <- read_xlsx(
  "input_data/cirulis_2026_archived_data/Total_mating_investment_G44.xlsx")
cirulis_dat2$Treatment <- factor(cirulis_dat2$Treatment)
cirulis_dat2$Population <- factor(cirulis_dat2$Population)
cirulis_dat2$`average mating time` <- scale(cirulis_dat2$`average mating time`)
```


#### Calculation of effect size

I used code provided by the original authors to run a binomial GLMM, where the response variable is whether the wild-type ("white") mating partner performed sucking behaviour after mating, evolution treatment and average mating time were fixed effects, and replicate population was a random factor. We these used `emmeans` to compute pairwise differences in sucking rate between the three treatments as log odds ratios, and converted the LORs to Cohen's $d$.

```{r}
model_sucking <- glmer(
  cbind(`White sucks`, matings-`White sucks`) ~ 
    `average mating time`+Treatment+(1|Treatment:Population), 
  data = cirulis_dat2, family = "binomial")

effects_sucking <- contrast(emmeans(model_sucking, "Treatment"), method = "pairwise") %>% 
  as.data.frame()
effects_sucking$estimate[1] <- effects_sucking$estimate[1] * -1
effects_sucking$Lower_95_CI <- effects_sucking$estimate - 1.96 * effects_sucking$SE
effects_sucking$Upper_95_CI <- effects_sucking$estimate + 1.96 * effects_sucking$SE

tibble(Study = "Cīrulis et al. (2026). Sex-limited experimental evolution, mating, and sexual conflict in a hermaphrodite. Zoo J Linnean Soc, 206, zlag036.",
       Trait = "Effect on mate's behaviour (female-limited lines)",
       Generations = 44,
       Organism = "Macrostomum lignano",
       Interpretation = "A positive effect size would indicate that worms carrying the female-limited genomic region induce more post-mating 'sucking' behaviour in their mates (they induced non-significantly less sucking, though).",
       Effect_size_source = "Model of the raw data",
       Estimate = effects_sucking$estimate[1],
       Var = effects_sucking$SE[1] ^ 2,
       Lower_95_CI = effects_sucking$Lower_95_CI[1],
       Upper_95_CI = effects_sucking$Upper_95_CI[1]) %>%
  write_csv("effect_sizes/cirulis2026_3.csv")


tibble(Study = "Cīrulis et al. (2026). Sex-limited experimental evolution, mating, and sexual conflict in a hermaphrodite. Zoo J Linnean Soc, 206, zlag036.",
       Trait = "Effect on mate's behaviour (male-limited lines)",
       Generations = 44,
       Organism = "Macrostomum lignano",
       Interpretation = "A positive effect size would indicate that worms carrying the male-limited genomic region induce less post-mating 'sucking' behaviour in their mates (they induced significantly less, though).",
       Effect_size_source = "Model of the raw data",
       Estimate = effects_sucking$estimate[2],
       Var = effects_sucking$SE[2] ^ 2,
       Lower_95_CI = effects_sucking$Lower_95_CI[2],
       Upper_95_CI = effects_sucking$Upper_95_CI[2]) %>% 
  write_csv("effect_sizes/cirulis2026_4.csv")
```


```{r}
stylet <- tibble(Treatment = c("C", "F", "M"),
       Mean = c(-0.009408707911216025,
                0.020585390642929746,
                -0.01103324481164552),
       SE = c(-0.0037006983377594165 - -0.009408707911216025,
              0.02691141274056589 - 0.020585390642929746,
              -0.007018622340251138 - -0.01103324481164552))

mes(m.1 = stylet$Mean[2], m.2 = stylet$Mean[1],
    sd.1 = stylet$SE[2] * sqrt(111), sd.2 = stylet$SE[1]  * sqrt(110),
    n.1 = 111, n.2 = 110, verbose = F) %>% 
  parse_es(study = "Cīrulis et al. (2026). Sex-limited experimental evolution, mating, and sexual conflict in a hermaphrodite. Zoo J Linnean Soc, 206, zlag036.",
           generations = 44,
           organism = "Macrostomum lignano",
           trait = "Stylet shape (female-limited lines)",
           interpretation = "The positive effect size indicates a different shape in the stylet (mating organ) of worms carrying the female-limited genomic region compared to controls. There was no a priori expectation of how shape would change, but since some change is predicted following removal of the male sex function we have coded this result as positive.") %>% 
  write_csv("effect_sizes/cirulis2026_5.csv")

mes(m.1 = stylet$Mean[1], m.2 = stylet$Mean[3],
    sd.1 = stylet$SE[1] * sqrt(110), sd.2 = stylet$SE[3]  * sqrt(114),
    n.1 = 110, n.2 = 114 , verbose = F) %>% 
  parse_es(study = "Cīrulis et al. (2026). Sex-limited experimental evolution, mating, and sexual conflict in a hermaphrodite. Zoo J Linnean Soc, 206, zlag036.",
           generations = 44,
           organism = "Macrostomum lignano",
           trait = "Stylet shape (male-limited lines)",
           interpretation = "A positive effect size would indicate a different shape in the stylet (mating organ) of worms carrying the arrying the male-limited genomic region compared to controls (with shape changing the opposite way to the change seen in the female-limited lines). There is no significant effect, however.") %>% 
  write_csv("effect_sizes/cirulis2026_6.csv")
```


## Melo-Gavin et al. (2026)

Melo-Gavin, C., Liu, M. J., & Agrawal, A. F. (2026). Local Adaptation to the Sex Environment: Reciprocal Sex-Limited Selection in Different Thermal Regimes. bioRxiv, 2026-01. https://doi.org/10.64898/2026.01.23.701411

### Information provided by the authors

The authors kindly re-ran a quasibinomial mixed model matching the one presented in the paper, then used `contrast(emmeans(model_fit, ~ Sex * Chromosome * Environment), method = "pairwise)"` to compute the effect sizes contrasting the chromosomes types (separately for each combination of sex and environment). The edited output is shown below (keeping only effect sizes that relate to fitness measures of one sex, measured in one environment). For example, the first row shows the difference in female fitness (`F`) between female and male-limited (`FL` and `ML`) chromosomes in the benign conditions (lines were both evolved in and measured in this environmental condition). The effect size estimate is a log odds ratio.

```{r}
df <- data.frame(
  Comparison = c(
    "FL F BENIGN - ML F BENIGN",
    "FL M BENIGN - ML M BENIGN",
    "FL F NOVEL - ML F NOVEL",
    "FL M NOVEL - ML M NOVEL"
  ),
  Estimate = c(0.3352, -0.1148, 0.3096, -0.7905),
  SE = c(0.136, 0.137, 0.136, 0.136)
)
```


## Effect size

Here we convert the log odds ratios to Cohen's d and adjust the effect size sign where needed.

```{r}
df <- df %>% 
  mutate(Cohens_d = Estimate * sqrt(3) / pi, 
         Cohens_d = Cohens_d * c(1,1,1,-1),
         Cohens_SE = SE * sqrt(3) / pi, 
         Cohens_Var = (SE^2) * sqrt(3) / pi, 
         Lower_95_CI = (Estimate * c(1,1,1,-1) - SE*1.96) * sqrt(3) / pi, 
         Upper_95_CI = (Estimate * c(1,1,1,-1) + SE*1.96) * sqrt(3) / pi) %>% 
  select(Comparison, Estimate = Cohens_d, Var = Cohens_Var, contains("95"))

tibble(Study = "Melo-Gavin et al. (2026) bioRxiv, https://doi.org/10.64898/2026.01.23.701411",
       Trait = c("Female fitness (benign env.)",
                 "Male fitness (benign env.)",
                 "Female fitness (novel env.)",
                 "Male fitness (novel env.)"),
       Generations = 15,
       Organism = "Drosophila melanogaster",
       Interpretation = c(
         "The positive effect size means that females carrying the female-limited chromosome had higher fitness than females with the male-limited chromosome, when evolved and measured under benign conditions.",
         "A positive effect size would mean that males carrying the male-limited chromosome had higher fitness than males with the female-limited chromosome, when evolved and measured under benign conditions (there is no difference though).",
         "The positive effect size means that females carrying the female-limited chromosome had higher fitness than females with the male-limited chromosome, when evolved and measured under novel conditions.",
         "The positive effect size means that males carrying the male-limited chromosome had higher fitness than males with the female-limited chromosome, when evolved and measured under novel conditions."),
       Effect_size_source = "Model of the raw data",
       Estimate = df$Estimate,
       Var = df$Var,
       Lower_95_CI = df$Lower_95_CI,
       Upper_95_CI = df$Upper_95_CI) %>% 
  write_csv("effect_sizes/MeloGavin2026.csv")
```

# Middle-class neighbourhood experiments

This section covers publications that used a sex-specific middle class neighbourhood to relax selection on one sex. In general, I predict reduced fitness for the unselected sex, increased fitness in the other sex (due to the spread of sexually antagonistic alleles benefitting the selected sex), and phenotypic evolution towards the typical trait values for the selected sex (e.g. in _Drosophila_, chromosomes that had undergone male-limited evolution in should encode smaller body size and slower development time). The effect size signs selected reflect these expectations.


## Radwan et al. (2004)

Radwan, J., Unrug, J., Snigórska, K., & Gawrońska, K. (2004). Effectiveness of sexual selection in preventing fitness deterioration in bulb mite populations under relaxed natural selection. _Journal of Evolutionary Biology_, 17(1), 94–99.

### Radwan's original data

The means and standard errors for each variable and treatment group were extracted from Table 1. Sample sizes for individual treatments are not given, so I divided the total sample size by 3 and rounded off to estimate them. 

```{r}
df <- tibble(
  Variable = rep(c("Viability",
                   "Female longevity",
                   "Female fecundity",
                   "Male fitness"),2),
  
  Treatment = c(rep("SS", 4), rep("C", 4)),
  
  mean = c(0.761, 23.74, 108.7, 0.543, 0.733, 25.35, 131.7, 0.596),
  SE   = c(0.102, 2.45, 4.47, 0.077, 0.066, 1.17, 9.51, 0.061),
  n = round(rep(c(161/3, 136/3, 139/3, 100/3), 2))
)

kable(df) %>% 
  kable_styling(full_width = F)
```

### Viability

```{r}
mes(m.2 = df$mean[df$Treatment=="SS" & df$Variable == "Viability"],
    m.1 = df$mean[df$Treatment=="C" & df$Variable == "Viability"],
    sd.2 = df$SE[df$Treatment=="SS" & df$Variable == "Viability"] * 
      sqrt(df$n[df$Treatment=="SS" & df$Variable == "Viability"]),
    sd.1 = df$SE[df$Treatment=="C" & df$Variable == "Viability"] * 
      sqrt(df$n[df$Treatment=="C" & df$Variable == "Viability"]),
    n.2 = df$n[df$Treatment=="SS" & df$Variable == "Viability"],
    n.1 = df$n[df$Treatment=="C" & df$Variable == "Viability"], verbose=F) %>% 
  parse_es(study = "Radwan et al. (2004) J. Evol. Biol. 17, 94–99.",
           generations = 11,
           organism = "Rhizoglyphus robini",
           trait = "Viability",
           interpretation = "A positive effect size would indicate lower viability in populations that experienced relaxed selection on females relative to those experiencing selection on both sexes (the difference is non-significant and in the opposite direction)."
          ) %>% 
  write_csv("effect_sizes/radwan2004_1.csv")
```

### Female longevity

```{r}
mes(m.2 = df$mean[df$Treatment=="SS" & df$Variable == "Female longevity"],
    m.1 = df$mean[df$Treatment=="C" & df$Variable == "Female longevity"],
    sd.2 = df$SE[df$Treatment=="SS" & df$Variable == "Female longevity"] * 
      sqrt(df$n[df$Treatment=="SS" & df$Variable == "Female longevity"]),
    sd.1 = df$SE[df$Treatment=="C" & df$Variable == "Female longevity"] * 
      sqrt(df$n[df$Treatment=="C" & df$Variable == "Female longevity"]),
    n.2 = df$n[df$Treatment=="SS" & df$Variable == "Female longevity"],
    n.1 = df$n[df$Treatment=="C" & df$Variable == "Female longevity"], verbose=F) %>% 
  parse_es(study = "Radwan et al. (2004) J. Evol. Biol. 17, 94–99.",
           generations = 11,
           organism = "Rhizoglyphus robini",
           trait = "Female longevity",
           interpretation = "A positive effect size would indicate reduced female longevity in populations that experienced relaxed selection on females relative to those experiencing selection on both sexes (the difference is non-significant and in the predicted direction).") %>% 
  write_csv("effect_sizes/radwan2004_2.csv")
```

### Female fecundity

```{r}
mes(m.2 = df$mean[df$Treatment=="SS" & df$Variable == "Female fecundity"],
    m.1 = df$mean[df$Treatment=="C" & df$Variable == "Female fecundity"],
    sd.2 = df$SE[df$Treatment=="SS" & df$Variable == "Female fecundity"] * 
      sqrt(df$n[df$Treatment=="SS" & df$Variable == "Female fecundity"]),
    sd.1 = df$SE[df$Treatment=="C" & df$Variable == "Female fecundity"] * 
      sqrt(df$n[df$Treatment=="C" & df$Variable == "Female fecundity"]),
    n.2 = df$n[df$Treatment=="SS" & df$Variable == "Female fecundity"],
    n.1 = df$n[df$Treatment=="C" & df$Variable == "Female fecundity"], verbose=F) %>% 
  parse_es(study = "Radwan et al. (2004) J. Evol. Biol. 17, 94–99.",
           generations = 11,
           organism = "Rhizoglyphus robini",
           trait = "Female fecundity",
           interpretation = "The positive effect size indicates reduced fecundity in populations that experienced relaxed selection on females relative to those experiencing selection on both sexes.") %>% 
  write_csv("effect_sizes/radwan2004_3.csv")
```

### Male fitness

```{r}
mes(m.1 = df$mean[df$Treatment=="SS" & df$Variable == "Male fitness"],
    m.2 = df$mean[df$Treatment=="C" & df$Variable == "Male fitness"],
    sd.1 = df$SE[df$Treatment=="SS" & df$Variable == "Male fitness"] * 
      sqrt(df$n[df$Treatment=="SS" & df$Variable == "Male fitness"]),
    sd.2 = df$SE[df$Treatment=="C" & df$Variable == "Male fitness"] * 
      sqrt(df$n[df$Treatment=="C" & df$Variable == "Male fitness"]),
    n.1 = df$n[df$Treatment=="SS" & df$Variable == "Male fitness"],
    n.2 = df$n[df$Treatment=="C" & df$Variable == "Male fitness"], verbose=F) %>% 
  parse_es(study = "Radwan et al. (2004) J. Evol. Biol. 17, 94–99.",
           generations = 11,
           organism = "Rhizoglyphus robini",
           trait = "Male fitness",
           interpretation = "A positive effect size would indicate increased fitness in populations that experienced relaxed selection on females relative to those experiencing selection on both sexes (the difference is non-significant and in the opposite direction).") %>% 
  write_csv("effect_sizes/radwan2004_4.csv")
```

## Morrow et al. (2008)

Morrow, E. H., Stewart, A. D., & Rice, W. R. (2008). Assessing the extent of genome-wide intralocus sexual conflict via experimentally enforced gender-limited selection. _Journal of Evolutionary Biology_, 21(4), 1046–1054.

### Morrow et al.'s original data

```{r}
df <- tibble(treatment = c("selected sex", "non-selected sex","selected sex", "non-selected sex"),
       trait = c("male fitness", "male fitness", "female fitness", "female fitness"),
       mean = c(9.761182759612618, 8.5105834691725, 38.60211669355156, 33.71168902702915),
       SE = c(10.07670917633522 - 9.761182759612618,
              8.812266276728353 - 8.5105834691725,
              38.99686561487718 - 38.60211669355156,
              33.985783792999435 - 33.71168902702915),
       N = c(60, 60, 60, 60))

kable(df) %>% 
  kable_styling(full_width = F)
```

### Male fitness effect size

```{r}
mes(m.1 = df$mean[df$trait == "male fitness" & df$treatment == "selected sex"],
    m.2 = df$mean[df$trait == "male fitness" & df$treatment == "non-selected sex"],
    sd.1 = df$SE[df$trait == "male fitness" & df$treatment == "selected sex"] * sqrt(60),
    sd.2 = df$SE[df$trait == "male fitness" & df$treatment == "non-selected sex"] * sqrt(60),
    n.1 = 60, n.2 = 60, verbose = F) %>% 
  parse_es(study = "Morrow et al. (2008) J. Evol. Biol. 21, 1046–1054",
           trait = "Male fitness", 
           generations = 26, 
           organism = "Drosophila melanogaster", 
           interpretation = "The positive effect size indicates that male fitness was higher in lines where selection was relaxed on females compared to lines where selection was relaxed on males.") %>% 
  write_csv("effect_sizes/morrow2008_1.csv")
```

### Female fitness effect size

```{r}
mes(m.1 = df$mean[df$trait == "female fitness" & df$treatment == "selected sex"],
    m.2 = df$mean[df$trait == "female fitness" & df$treatment == "non-selected sex"],
    sd.1 = df$SE[df$trait == "female fitness" & df$treatment == "selected sex"] * sqrt(60),
    sd.2 = df$SE[df$trait == "female fitness" & df$treatment == "non-selected sex"] * sqrt(60),
    n.1 = 60, n.2 = 60, verbose = F) %>% 
  parse_es(study = "Morrow et al. (2008) J. Evol. Biol. 21, 1046–1054",
           trait = "Female fitness", 
           generations = 26, 
           organism = "Drosophila melanogaster", 
           interpretation = "The positive effect size indicates that female fitness was higher in lines where selection was relaxed on males compared to lines where selection was relaxed on females.") %>% 
  write_csv("effect_sizes/morrow2008_2.csv")
```

## Bonel et al. (2018)

Bonel, N., Noël, E., Janicke, T., Sartori, K., Chapuis, E., Segard, A., Meconcelli, S., Pelissie, B., Sarda, V. & David, P. (2018). Asymmetric evolutionary responses to sex‐specific selection in a hermaphrodite. _Evolution_, 72(10), 2181-2201.

The raw data were obtained from the Dryad repository. 

### Female fecundity 

I used the generation 20 data for convenience (the variables are named in French, and there are many files with many columns each in the Dryad archive, so I did not attempt a more complex analysis). I fit a linear mixed model of the fecundity data with block, family ID, and replicate evolution line as random factors. 

```{r message=F, warning=F}
bonel1 <- read.csv2("input_data/Bonel_2018_snails_dryad_data/G20.female.rep.function.csv", check.names = F) %>% as_tibble()

# bonel2 <- read.csv("input_data/Bonel_2018_snails_dryad_data/G35.reproductive.traits.csv", check.names = F) %>% as_tibble() %>% rename(treatment = linetype, replicate = line) %>% pull(eggs3)


names(bonel1)[43] <- "female_fecundity" # original name: "oeufs maternel total corrigés"
names(bonel1)[3] <- "treatment" # original name: lignée
names(bonel1)[4] <- "replicate" # original name: réplicat

bonel1 <- bonel1 %>%
  mutate(female_fecundity = scale(female_fecundity),
         treatment = factor(treatment),
         treatment = relevel(treatment, ref = "M"))

fecundity_model <- lmer(
  female_fecundity ~ treatment + (1 | BLOC) + (1 | famille) + (1 | replicate), 
  data = bonel1)
coef <- coefficients(summary(fecundity_model)) %>% 
  as.data.frame()

tibble(
  Estimate = coef$Estimate[3],
  Var = coef[3,2] ^ 2,
  Lower_95_CI = Estimate - coef[3,2] * 1.96,
  Upper_95_CI = Estimate + coef[3,2] * 1.96) %>% 
    mutate(Study = "Bonel et al. (2018) Evolution, 72, 2181-2201", 
           Trait = "Female fecundity", 
           Generations = "20",
           Organism = "Physa acuta",
           Interpretation = "A positive effect size would indicate that female fecundity was higher in lines where selection was relaxed on males compared to lines where selection was relaxed on females (it's non-significantly different in the opposite direction).",
           Effect_size_source = "Calculated from the raw data using a model",
           .before = 1) %>% 
  write_csv("effect_sizes/bonel2018_1.csv")
```

### Siring success

```{r}
bonel2 <- read.csv2("input_data/Bonel_2018_snails_dryad_data/G20.male.rep.function.csv", 
                    check.names = F) 

names(bonel2)[7] <- "treatment"
bonel2$treatment[bonel2$treatment == "T"] <- "C"
names(bonel2)[5] <- "replicate"

# There are columns called survivantsPmèretotal and survivantsAmèretotal
# I read this as surviving offspring from the focal (P) father and alternative/albino (A) father
# I used these as the 2-column response variable for binomial GLM predicting siring success
resp <- cbind(bonel2[,47], bonel2[,48])


bonel2 <- bonel2 %>%
  mutate(treatment = factor(treatment),
         treatment = relevel(treatment, ref = "M"))

# There is no variance associated with replicate line, so I removed it. It's a bit confounded with BLOC
siring_model <- glmer(resp ~ treatment + (1 | BLOC) + (1 | idfamille_focal), data = bonel2, family = binomial)
coef <- coefficients(summary(siring_model)) %>% 
  as.data.frame()


tibble(
  Estimate = -1 * coef$Estimate[3] * sqrt(3) / pi,
  Var = (coef[3,2] ^ 2) * sqrt(3) / pi,
  Lower_95_CI = (-1 * coef$Estimate[3] - coef[3,2]*1.96) * sqrt(3) / pi,
  Upper_95_CI = (-1 * coef$Estimate[3] + coef[3,2]*1.96) * sqrt(3) / pi) %>% 
    mutate(Study = "Bonel et al. (2018) Evolution, 72, 2181-2201", 
           Trait = "Male siring success", 
           Generations = "20",
           Organism = "Physa acuta",
           Interpretation = "The positive effect size indicates that siring success as the male parent was higher in lines where selection was relaxed on females compared to lines where selection was relaxed on males.",
           Effect_size_source = "Calculated from the raw data using a model",
           .before = 1) %>% 
  write_csv("effect_sizes/bonel2018_2.csv")
```

